---
title: "LandWeb v3: Overview"
subtitle: "Historic Natural Range of Variation modelling in the western boreal --- and how to read this report suite"
author: "Alex M. Chubaty"
date: today
format:
  pdf:
    pdf-engine: pdflatex
    toc: true
    toc-depth: 2
    number-sections: true
    fontsize: 11pt
fig-width: 7
fig-height: 6.5
bibliography:
  - ../citations/references.bib
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(data.table)
library(ggplot2)
library(kableExtra)
```

\newpage

::: {.callout-warning}
**Draft report: pending human review**

Portions of this document were drafted with the help of generative AI (large language models) and
have not yet been fully reviewed by the co-authors.
Content may be incomplete, inaccurate, or misattributed.
This notice will be removed once a complete review has been carried out.
:::

# What LandWeb is

LandWeb is a data-driven modelling configuration for simulating the **Historic (Natural) Range of
Variation (H/NRV)** of forest condition across most of the western Canadian boreal forest
(~3 million km^2^). Rather than a single model, it couples two well-established simulation models
inside the `SpaDES` platform [@R-SpaDES.core] --- a workflow embodiment of the PERFICT approach to
predictive ecology [@McIntire:2022; @Barros2023]:

- **vegetation dynamics** --- an implementation of the LANDIS-II Biomass Succession model
  [@SchellerMladenoff2004; @Scheller2007], simulating growth, mortality, dispersal, and
  post-disturbance regeneration of tree-species cohorts on each pixel; and
- **wildfire** --- an implementation of the LandMine landscape-disturbance model
  [@Andison:1996; @Andison:1998], igniting and spreading fires at rates set by a long-term
  historic fire-cycle (LTHFC) map.

Run over many centuries and many stochastic replicates, the simulation produces a *distribution* of
plausible pre-industrial landscape states --- the natural range of variation --- against which the
current landscape can be compared. The pre-industrial fire regime that anchors the fire model comes
from a long-term historic fire-cycle synthesis for the western boreal [@Andison:2019].

# The v3 update at a glance

v3 is primarily a **rebuild of the modelling infrastructure and a refresh of the input data**, not a
change of ecological content. The main changes (each documented in its own report, below):

- **Inputs.** Tree-species / biomass / age initial conditions move from the Beaudoin *k*NN 2001 maps
  [@BeaudoinEtAl2014] to SCANFI 2020 [@scanfi-v2] (report 01); the fire-cycle map is updated to
  **v10**, incorporating the June-2026 Northwest Alberta work [@peterson-nwab-data] (report 02).
- **Orchestration.** The manual `SpaDES` script chain is replaced by a **`targets` pipeline** that
  owns caching and dispatch, with the `SpaDES` modules retained as the simulation engine (report 03).
- **Scope.** Simulations are run and reported **per forest-management area (FMA)** rather than as one
  monolithic run.
- **Post-processing.** Three v2 summary modules are consolidated into the new `NRV_summary` module
  plus the `nrvtools` package (report 03).

# Modelling workflow

```{r}
#| label: fig-workflow
#| fig-cap: "The LandWeb v3 modelling workflow: input data feed a data-preparation stage, a main simulation, and a post-processing (NRV analyses) stage that produces the summary outputs. In v3 the post-processing modules NRV_summary and burnSummaries replace the v2 trio timeSinceFire / LandWeb_output / LandWeb_summary."
wf <- data.frame(
  y = 5:1,
  title = c("Input data", "Data preparation", "Main simulation",
    "Post-processing (NRV analyses)", "Model outputs"),
  body = c(
    "forest cover  |  vegetation parameters  |  fire history  |  study area & reporting polygons",
    "LandWeb_preamble  |  Biomass_speciesData  |  Biomass_speciesParameters  |  Biomass_speciesFactorial  |  Biomass_borealDataPrep",
    "Biomass_core  |  Biomass_regeneration  |  LandMine",
    "NRV_summary  |  burnSummaries",
    "time-since-fire & stand-age maps  |  leading-vegetation, age-class & large-patch summaries  |  fire-size distributions"
  ),
  fill = c("#dbe9f6", "#fbf1cf", "#e3f1e3", "#ece5f4", "#f6ead4"),
  stringsAsFactors = FALSE
)
wrapl <- function(s, w = 62) vapply(s, function(x) paste(strwrap(x, width = w), collapse = "\n"), character(1))
wf$body_w <- wrapl(wf$body)
arrows <- data.frame(ytop = c(5, 4, 3, 2))

ggplot(wf) +
  geom_rect(aes(xmin = 0.2, xmax = 9.8, ymin = y - 0.38, ymax = y + 0.38, fill = fill),
    color = "grey55", linewidth = 0.4) +
  geom_text(aes(x = 5, y = y + 0.24, label = title), fontface = "bold", size = 4) +
  geom_text(aes(x = 5, y = y - 0.10, label = body_w), size = 2.7, lineheight = 0.95) +
  geom_segment(data = arrows, aes(x = 5, xend = 5, y = ytop - 0.38, yend = ytop - 0.62),
    arrow = arrow(length = unit(0.18, "cm"), type = "closed"), color = "grey35", linewidth = 0.6) +
  scale_fill_identity() +
  coord_cartesian(xlim = c(0, 10), ylim = c(0.5, 5.6), clip = "off") +
  theme_void()
```

The `targets` pipeline runs the data-preparation stage once per study area (caching the
study-area-independent species trait factorial), then crosses the prepared data with a set of
stochastic replicates for the main simulation, and reduces the replicates into the NRV summaries.
The generic `SpaDES`-to-`targets` glue lives in the owned `SpaDES.targets` package; heavy stages are
dispatched across a small compute cluster via `crew.ssh` (report 03).

# Study areas

v3 treats the **forest-management area (FMA)** as the unit of simulation and reporting: each area is
parameterized, simulated, and summarized on its own, rather than as part of a single continental run
(the v2 `LandWeb_full` run is dropped). This keeps per-area updates from invalidating unrelated
results and keeps peak memory bounded. The FMAs span the western boreal across British Columbia,
Alberta, Saskatchewan, Manitoba, and the Northwest Territories; the historical per-FMA client
deliverables in `reports/2020/` indicate the scope. The v3 pipeline was first validated end-to-end
on the **Spray Lake** study area (SW Alberta).

# This report suite

The v3 documentation is organized as a numbered suite (this overview is report 00), plus the
separate user manual. Reports are rendered from the pipeline so their figures and numbers stay
consistent with the model.

```{r}
#| label: tbl-suite
#| tbl-cap: "The LandWeb v3 report suite."
suite <- data.table(
  Report = c("00", "01", "02", "03", "04", "05", "06", "07", "Manual"),
  Contents = c(
    "Overview and rationale (this document); modelling workflow; report index.",
    "Study-area vegetation: SCANFI composition, and the kNN (v2) to SCANFI (v3) initial-condition change.",
    "Fire-model inputs: the LandMine LTHFC v10 (2026 NW Alberta) update vs the previous v8c / Spray Lake layers.",
    "Code and module changes v2 to v3: per-module/package history; SpaDES framework breakage and the targets switch.",
    "Output differences v2 vs v3: how the simulated NRV and burn summaries change (planned).",
    "Per-study-area summary: the templated NRV report rendered per FMA --- the client-facing deliverable (planned).",
    "Provenance: the input/output manifest and reproducibility standards (planned).",
    "Best-practices alignment vs forest-landscape-model calibration guidance (planned, optional).",
    "User guide / manual (migrating from bookdown to a Quarto book), tracked by issue #128."
  )
)
knitr::kable(suite, align = "ll", booktabs = TRUE) |>
  kable_styling(font_size = 8.5, latex_options = "hold_position") |>
  column_spec(1, width = "3.5em") |>
  column_spec(2, width = "38em")
```

\newpage

# References {.unnumbered}

::: {#refs}
:::

\newpage

# Provenance {.unnumbered}

- **Source:** this overview is narrative; the workflow figure (@fig-workflow) is a fixed schematic
  built in `ggplot2` from the current v3 module set (cf. the DiagrammeR source in
  `R/model_workflow.R`).
- **Method:** rendered by the gated `landweb.extended_analyses` factory as a
  `tar_quarto()` target on the control node; no pipeline data dependencies.

{{< include _provenance-appendix.qmd >}}
