---
title: "LandWeb v3: Overview"
subtitle: "Historic Natural Range of Variation modelling in the western boreal --- and how to read this report suite"
author: "Alex M. Chubaty"
date: today
format:
  pdf:
    pdf-engine: pdflatex
    toc: true
    toc-depth: 2
    number-sections: true
    fontsize: 11pt
fig-width: 7
fig-height: 6.5
bibliography:
  - ../citations/references.bib
csl: ../citations/ecology-letters.csl
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(data.table)
library(ggplot2)
library(kableExtra)
library(sf)

## study-area group reference artifacts (study-area-independent -> outputs/_reference/);
## regenerate with scripts/make_sa_reference.R. Paths tried from both the project root and
## the report dir so the report also renders standalone.
.ref <- function(f) {
  cand <- file.path(c("outputs/_reference", "../outputs/_reference"), f)
  cand[file.exists(cand)][1]
}
sa_csv <- .ref("studyAreaGroups.csv")
sa_gpkg <- .ref("studyAreaGroups.gpkg")
sa_tab <- if (!is.na(sa_csv)) fread(sa_csv) else NULL
sa_report_png <- .ref("studyAreaGroups_report.png")
sa_facet_pngs <- Filter(Negate(is.na), lapply(1:3, function(i) .ref(sprintf("studyAreaGroups_facets_p%d.png", i))))
```

\newpage

::: {.callout-warning}
**Draft report: pending human review**

Portions of this document were drafted with the help of generative AI (large language models) and
have not yet been fully reviewed by the co-authors.
Content may be incomplete, inaccurate, or misattributed.
This notice will be removed once a complete review has been carried out.
:::

# What LandWeb is

LandWeb is a data-driven modelling configuration for simulating the **Historic (Natural) Range of
Variation (H/NRV)** of forest condition across most of the western Canadian boreal forest
(~3 million km^2^). Rather than a single model, it couples two well-established simulation models
inside the `SpaDES` platform [@R-SpaDES.core] --- a workflow embodiment of the PERFICT approach to
predictive ecology [@McIntire:2022; @Barros2023]:

- **vegetation dynamics** --- an implementation of the LANDIS-II Biomass Succession model
  [@SchellerMladenoff2004; @Scheller2007], simulating growth, mortality, dispersal, and
  post-disturbance regeneration of tree-species cohorts on each pixel; and
- **wildfire** --- an implementation of the LandMine landscape-disturbance model
  [@Andison:1996; @Andison:1998], igniting and spreading fires at rates set by a long-term
  historic fire-cycle (LTHFC) map.

Run over many centuries and many stochastic replicates, the simulation produces a *distribution* of
plausible pre-industrial landscape states --- the natural range of variation --- against which the
current landscape can be compared. The pre-industrial fire regime that anchors the fire model comes
from a long-term historic fire-cycle synthesis for the western boreal [@Andison:2019].

# The v3 update at a glance

v3 is primarily a **rebuild of the modelling infrastructure and a refresh of the input data**, not a
change of ecological content. The main changes (each documented in its own report, below):

- **Inputs.** Tree-species / biomass / age initial conditions move from the Beaudoin *k*NN 2001 maps
  [@BeaudoinEtAl2014] to SCANFI 2020 [@scanfi-v2] (report 01); the fire-cycle map is updated to
  **v10**, incorporating the June-2026 Northwest Alberta work [@peterson-nwab-data] (report 02).
- **Orchestration.** The manual `SpaDES` script chain is replaced by a **`targets` pipeline** that
  owns caching and dispatch, with the `SpaDES` modules retained as the simulation engine (report 03).
- **Scope.** Simulations are run per **ecoregion-based study-area group** --- ecologically coherent
  bundles of adjacent forest tenures --- rather than as one monolithic run, with the per-tenure
  breakdown recovered through the reporting polygons.
- **Post-processing.** Three v2 summary modules are consolidated into the new `NRV_summary` module
  plus the `nrvtools` package (report 03).

# Modelling workflow

```{r}
#| label: fig-workflow
#| fig-cap: "The LandWeb v3 modelling workflow: input data feed a data-preparation stage, a main simulation, and a post-processing (NRV analyses) stage that produces the summary outputs. In v3 the post-processing modules NRV_summary and burnSummaries replace the v2 trio timeSinceFire / LandWeb_output / LandWeb_summary."
wf <- data.frame(
  y = 5:1,
  title = c("Input data", "Data preparation", "Main simulation",
    "Post-processing (NRV analyses)", "Model outputs"),
  body = c(
    "forest cover  |  vegetation parameters  |  fire history  |  study area & reporting polygons",
    "LandWeb_preamble  |  Biomass_speciesData  |  Biomass_speciesParameters  |  Biomass_speciesFactorial  |  Biomass_borealDataPrep",
    "Biomass_core  |  Biomass_regeneration  |  LandMine",
    "NRV_summary  |  burnSummaries",
    "time-since-fire & stand-age maps  |  leading-vegetation, age-class & large-patch summaries  |  fire-size distributions"
  ),
  fill = c("#dbe9f6", "#fbf1cf", "#e3f1e3", "#ece5f4", "#f6ead4"),
  stringsAsFactors = FALSE
)
wrapl <- function(s, w = 62) vapply(s, function(x) paste(strwrap(x, width = w), collapse = "\n"), character(1))
wf$body_w <- wrapl(wf$body)
arrows <- data.frame(ytop = c(5, 4, 3, 2))

ggplot(wf) +
  geom_rect(aes(xmin = 0.2, xmax = 9.8, ymin = y - 0.38, ymax = y + 0.38, fill = fill),
    color = "grey55", linewidth = 0.4) +
  geom_text(aes(x = 5, y = y + 0.24, label = title), fontface = "bold", size = 4) +
  geom_text(aes(x = 5, y = y - 0.10, label = body_w), size = 2.7, lineheight = 0.95) +
  geom_segment(data = arrows, aes(x = 5, xend = 5, y = ytop - 0.38, yend = ytop - 0.62),
    arrow = arrow(length = unit(0.18, "cm"), type = "closed"), color = "grey35", linewidth = 0.6) +
  scale_fill_identity() +
  coord_cartesian(xlim = c(0, 10), ylim = c(0.5, 5.6), clip = "off") +
  theme_void()
```

The `targets` pipeline runs the data-preparation stage once per study area (caching the
study-area-independent species trait factorial), then crosses the prepared data with a set of
stochastic replicates for the main simulation, and reduces the replicates into the NRV summaries.
The generic `SpaDES`-to-`targets` glue lives in the owned `SpaDES.targets` package; heavy stages are
dispatched across a small compute cluster via `crew.ssh` (report 03).

# Study areas

v3 runs the model over a set of **ecoregion-based study-area groups** rather than one monolithic
continental run (the v2 `LandWeb_full` run is dropped) or one run per forest tenure. Each group
bundles forest tenures that are both **geographically adjacent** and **in the same ecological
region**; it is parameterized, simulated, and summarized as a unit, and the per-tenure numbers
partners need come from the *reporting polygons* (next section), not from running each tenure
separately.

## Why groups, and why ecoregions

Two problems drove the design, and neither is solved by running tenures individually.

**Small tenures cannot be parameterized alone.** Species traits are fitted from permanent sample
plot (PSP) data cropped to the study area. Run a small tenure by itself and its PSP catchment is too
small --- per-species samples fall below the 50-plot minimum and the trait fit fails outright. This
is not hypothetical: a solo run of a ~2,700 km^2^ Alberta tenure dies in
`Biomass_speciesParameters` for exactly this reason. Small tenures therefore have to be run *inside*
a larger area.

**But a larger area must be ecologically coherent.** Species traits legitimately differ between
ecological regions, so a PSP pool that spans very different ecology produces fits that are wrong
everywhere in it. Bundling tenures purely by proximity --- or by corporate ownership --- would do
exactly that.

Two ecological units are therefore involved, at different levels, and it is worth keeping them
apart:

- **Group membership** --- which tenures are simulated together, and what the group is called ---
  is set by **ecoregion**: each tenure is assigned to the ecoregion it mostly falls in, and the
  group takes that ecoregion's name (`studyAreaCrosswalk(eco_field = "REGION_NAM")`). All 18 group
  names are ecoregion names.
- **The PSP catchment** used to fit species traits (`studyAreaANPP`) is **coarser still**: it is
  the set of **ecoprovinces** the study area intersects. A single ecoregion does not supply enough
  plots, which is exactly the failure that motivated the regrouping.

So the groups are ecoregion-*defined* but their trait data are pooled at the ecoprovince level ---
deliberately, to trade a little ecological specificity for a sample size that supports a fit.

A third consideration settled *how* membership is defined. The former study-area registry was keyed
to companies and referenced tenure polygons by ID, and neither survives contact with updated
boundary data: tenure IDs are re-issued between versions, and tenures change hands, so a polygon
labelled for one licensee in an older file may belong to another today. Group membership is
therefore established by **verified spatial overlap against the current boundary layer**, never by
inherited labels --- geography and ecology are stable, corporate identity is not. The grouping is
recorded in an explicit crosswalk (`LandWebUtils::studyAreaCrosswalk()`), which retains `company`
and `province` columns so a licensee-driven rerun still maps cleanly onto the groups holding their
tenures.

The grouping layer is a parameter, not a constant: a coarser layer (ecoprovinces) yields fewer,
larger groups and a finer one (ecodistricts) subdivides any group that proves too large to run.

## The groups

```{r}
#| label: fig-sa-groups
#| fig-cap: "LandWeb v3 study-area groups. Each bold outline is one study area --- the union of its member forest tenures, whose individual boundaries are drawn within it. Every group has a unique colour + hatch combination, and neighbouring groups always differ in colour; numbers key to @tbl-sa-groups. Provinces and the ecoregions the groups are named for (dashed) are shown for context."
#| out-width: "100%"
## The figure is pre-rendered by scripts/make_sa_reference.R rather than rebuilt here: it needs the
## FMA layer, the crosswalk and a GADM download, and rendering it in-report would duplicate ~40
## lines of plotting code that could then drift from the standalone/shareable copy.
if (!is.na(sa_report_png)) {
  knitr::include_graphics(normalizePath(sa_report_png))
} else {
  knitr::asis_output("*(study-area map not found --- run `scripts/make_sa_reference.R`)*")
}
```

```{r}
#| label: tbl-sa-groups
#| tbl-cap: "Study-area groups and their member forest tenures. `Mpix` is the approximate simulated pixel count at 240 m resolution --- the practical size measure for run planning."
if (!is.null(sa_tab)) {
  out <- sa_tab[, .(`#` = id, `Study area group` = group, Prov = province, N = n_members,
                    `Member tenures` = members,
                    `Area (km^2^)` = format(area_km2, big.mark = ","), Mpix = mpix)]
  kbl(out, booktabs = TRUE, longtable = TRUE, align = "rllrlrr") |>
    kable_styling(latex_options = c("striped", "repeat_header"), font_size = 7) |>
    column_spec(5, width = "16em")
} else {
  knitr::asis_output("*(study-area lookup not found --- run `scripts/make_sa_reference.R`)*")
}
```

```{r}
#| label: fig-sa-facets
#| fig-cap: "Member forest tenures (blue) of each study-area group, one panel per group. All panels share a single map scale, so group sizes are directly comparable both within and between pages. The tan area is the ecoregion the group is named for, which is what defines its membership, so it is clear where a group's tenures spill beyond it; dashed lines are ecoprovince boundaries --- the coarser unit the PSP trait pool (`studyAreaANPP`) is drawn from. Tenure labels are the curated short names used throughout the reporting, and panel numbers match @fig-sa-groups and @tbl-sa-groups."
#| out-width: "100%"
if (length(sa_facet_pngs)) {
  knitr::include_graphics(normalizePath(unlist(sa_facet_pngs)))
} else {
  knitr::asis_output("*(study-area facet plate not found --- run `scripts/make_sa_reference.R`)*")
}
```

`r if (!is.null(sa_tab)) sprintf("The %d groups cover %s km^2^ across %d member tenures in British Columbia, Alberta, Saskatchewan, Manitoba, Ontario, and the Northwest Territories. They span roughly %s-fold in area (%s--%s km^2^), so run cost varies accordingly.", nrow(sa_tab), format(sum(sa_tab$area_km2), big.mark = ","), sum(sa_tab$n_members), round(max(sa_tab$area_km2) / min(sa_tab$area_km2)), format(min(sa_tab$area_km2), big.mark = ","), format(max(sa_tab$area_km2), big.mark = ","))`
`r if (!is.null(sa_tab)) sprintf("The v3 pipeline was first validated end-to-end on **WesternAlbertaUpland** (group %d), whose %d member tenures make it the densest-tenure group and therefore the most demanding test of the per-tenure reporting described next.", sa_tab[group == "WesternAlbertaUpland", id], sa_tab[group == "WesternAlbertaUpland", n_members])`

# Reporting polygons

Results are summarized over **reporting polygons**: the tenure boundaries themselves, a set of
sub-region layers (caribou ranges, Alberta Natural Subregions, BC biogeoclimatic zones, national
ecozones/ecoregions, NWT ecoregions, parks), and --- restored in v3 --- the **crossings** of each
tenure with each sub-region layer, so a partner sees their own share of each sub-region rather
than a study-area-wide total. Crossed units are named `<tenure> <layer>` (e.g. `SprayLake ANSR`),
and the sub-regions within them `<sub-region> <tenure>`. Crossing happens *before* the
name-grouping that aggregates a sub-region's disjoint parts, so two tenures' shares of the same
sub-region are never pooled.

::: {.callout-note}
## Caveats on the caribou range reporting

**Alberta range granularity --- REVISIT.** Alberta caribou units are reported at the **range**
level (`LOCALRANGE`, 15 ranges), matching v2. The source layer also carries a finer **sub-range**
attribute (`SUBUNIT`, 26 units) which would, for example, report Algar, Egg-Pony, Wiau, Agnes,
Bohn, Christina, and Wandering separately instead of together as East Side Athabasca. Range was
chosen for continuity with v2 reporting, but the appropriate granularity should be confirmed with
partners before the v3 reporting is finalized.

**Northwest Territories ranges --- interim source.** The current combined caribou layer contains
no NWT ranges, so NWT caribou reporting is supplemented with the single `Northwest Territories
(NWT)` range polygon from the layer used before that update. NWT tenures therefore get one
whole-territory caribou unit rather than a sub-region breakdown. A replacement has been identified
--- the Government of the Northwest Territories' *Boreal Caribou Range Planning Regions*, which
subdivides the NT1 range into six named regions (Southern NWT, Sahtú, Wek'èezhìi, Gwich'in,
Inuvialuit, Yukon) --- pending agreement on folding it into the maintained caribou layer.
:::

# This report suite

The v3 documentation is organized as a numbered suite (this overview is report 00), plus the
separate user manual. Reports are rendered from the pipeline so their figures and numbers stay
consistent with the model.

```{r}
#| label: tbl-suite
#| tbl-cap: "The LandWeb v3 report suite."
suite <- data.table(
  Report = c("00", "01", "02", "03", "04", "05", "06", "07", "Manual"),
  Contents = c(
    "Overview and rationale (this document); modelling workflow; report index.",
    "Study-area vegetation: SCANFI composition, and the kNN (v2) to SCANFI (v3) initial-condition change.",
    "Fire-model inputs: the LandMine LTHFC v10 (2026 NW Alberta) update vs the previous v8c / Spray Lake layers.",
    "Code and module changes v2 to v3: per-module/package history; SpaDES framework breakage and the targets switch.",
    "Output differences v2 vs v3: how the simulated NRV and burn summaries change (planned).",
    "Per-study-area summary: the templated NRV report rendered per FMA --- the client-facing deliverable (planned).",
    "Provenance: the input/output manifest and reproducibility standards (planned).",
    "Best-practices alignment vs forest-landscape-model calibration guidance (planned, optional).",
    "User guide / manual (migrating from bookdown to a Quarto book), tracked by issue #128."
  )
)
knitr::kable(suite, align = "ll", booktabs = TRUE) |>
  kable_styling(font_size = 8.5, latex_options = "hold_position") |>
  column_spec(1, width = "3.5em") |>
  column_spec(2, width = "38em")
```

\newpage

# References {.unnumbered}

::: {#refs}
:::

\newpage

# Provenance {.unnumbered}

- **Source:** this overview is narrative; the workflow figure (@fig-workflow) is a fixed schematic
  built in `ggplot2` from the current v3 module set (cf. the DiagrammeR source in
  `R/model_workflow.R`).
- **Method:** rendered by the gated `landweb.extended_analyses` factory as a
  `tar_quarto()` target on the control node; no pipeline data dependencies.

{{< include _provenance-appendix.qmd >}}
