---
title: "LandWeb Study-Area Vegetation"
subtitle: "SCANFI species composition over the LandWeb study area"
author: "Alex M. Chubaty"
date: today
format:
  pdf:
    pdf-engine: pdflatex
    toc: true
    toc-depth: 2
    number-sections: true
    fontsize: 11pt
fig-width: 8
fig-height: 6
bibliography:
  - ../citations/references.bib
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(targets)
library(data.table)
library(ggplot2)
library(kableExtra)

## display absolute pipeline paths as short project-relative paths
relpath <- function(p) sub("^.*/((?:inputs|outputs)/.*)$", "\\1", p, perl = TRUE)

scanfi_csv <- tar_read(scanfi_veg_summary_csv)
veg <- fread(scanfi_csv)
area_cells <- veg$domain_forested_cells[1]
area_km2 <- veg$domain_area_km2[1]
suspects <- c("TSUG_HET", "THUJ_PLI", "PINU_CON_CON")
veg[, suspect := species %in% suspects]
pct <- function(code) veg[species == code, pct_forested]
```

\newpage

::: {.callout-warning}
**Draft report: pending human review**

Portions of this document were drafted with the help of generative AI (large language models) and have not yet been fully reviewed by the co-authors.
Content may be incomplete, inaccurate, or misattributed.
This notice will be removed once a complete review has been carried out.
:::

# Overview

This report summarizes the SCANFI v2 (2020) [@scanfi-v2] per-species
**crown-closure prevalence** over the **full LandWeb study area**
(`r format(round(area_km2), big.mark = ",")` km^2^,
`r format(area_cells, big.mark = ",")` forested 30 m cells). For each SCANFI
source species that maps to a simulated LandWeb species it reports the number
(and %) of forested study-area cells carrying *any* cover.

Its primary purpose is to document the evidence behind the
**`THUJ_PLI`/`TSUG_HET` → `Abie_spp` merge** in `LandWeb_preamble`: western
redcedar and western hemlock are a *localized* SCANFI over-attribution in the
SW Alberta Rockies, not a study-area-wide signal.

# Study area composition

```{r}
#| label: tbl-composition
#| tbl-cap: "SCANFI species prevalence over the LandWeb study area, by cells with any cover (sorted). `LandWeb species` is the simulated species each SCANFI code merges into."
knitr::kable(
  veg[, .(
    `SCANFI code` = species,
    `Common name` = common_name,
    `LandWeb species` = landweb_species,
    `Cells (cover > 0)` = format(n_cover_gt0, big.mark = ","),
    `% of study area` = sprintf("%.3f", pct_forested)
  )],
  align = "lllrr",
  booktabs = TRUE
) |>
  kable_styling(font_size = 8, latex_options = "hold_position")
```

```{r}
#| label: fig-composition
#| fig-cap: "Prevalence (% of forested study-area cells with cover > 0) by SCANFI species; coastal-attribution suspects in red."
ggplot(veg, aes(reorder(common_name, pct_forested), pct_forested, fill = suspect)) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(values = c(`FALSE` = "grey40", `TRUE` = "firebrick"), guide = "none") +
  labs(x = NULL, y = "% of forested study-area cells with cover > 0") +
  theme_minimal(base_size = 12)
```

The composition is a textbook western-boreal signature --- black spruce
dominant, then tamarack, aspen, jack pine, and white spruce.

# Coastal over-attribution (the merge rationale)

```{r}
#| label: tbl-suspects
#| tbl-cap: "Coastal-attribution suspects: study-area prevalence and substantial-cover (>= 10%) extent."
knitr::kable(
  veg[species %in% suspects][
    order(-pct_forested),
    .(
      `SCANFI code` = species,
      `Common name` = common_name,
      `% study area (cover > 0)` = sprintf("%.3f", pct_forested),
      `% study area (>= 10%)` = sprintf("%.4f", pct_ge10),
      `Max cover %` = cover_max
    )
  ],
  align = "llrrr",
  booktabs = TRUE
) |>
  kable_styling(font_size = 8, latex_options = "hold_position")
```

Western hemlock (`TSUG_HET`) covers only **`r sprintf("%.2f", pct("TSUG_HET"))`%**
of the LandWeb study area, and western redcedar (`THUJ_PLI`)
**`r sprintf("%.2f", pct("THUJ_PLI"))`%** --- rounding errors at the study-area
scale. By contrast, a stand-alone count over the **Spray Lake FMA** (SW Alberta
Rockies) put western hemlock on **~25%** of pixels: a ~175× concentration in the
SW corner, exactly where SCANFI bleeds coastal species across the BC boundary.
Where these species do appear they can reach high cover (up to
`r veg[species == "TSUG_HET", cover_max]`% hemlock), but the footprint is tiny
and localized.

Because both fold into `Abie_spp` (the closest shade-tolerant softwood), the
merge:

- removes a stand-alone species that has no usable known-age / non-zero-biomass
  cohorts in the SW --- the cause of the `Biomass_borealDataPrep` non-conformable
  ("code 235") crash; and
- relabels only ~0.3% of the study area's cover, leaving the rest untouched.

`PINU_CON_CON` (shore pine, the coastal *Pinus contorta* variety) is effectively
absent across the study area
(`r format(veg[species == "PINU_CON_CON", n_cover_gt0], big.mark = ",")` cells),
so it is left merged into `Pinu_spp`, where --- like all pines --- it is harmless.

\newpage

# References {.unnumbered}

::: {#refs}
:::

\newpage

# Provenance {.unnumbered}

- **Source:** SCANFI v2 species crown-closure layers [@scanfi-v2], derived via
  the SCANFI approach [@scanfi-methods]
  (`SCANFI_spsCC_<code>_2020_v2_20260119.tif`), masked to the LandWeb study area
  (the LTHFC v10 union, `landweb_ltfc_v10`).
- **Data file:** `` `r relpath(scanfi_csv)` ``
- **Method:** `R/scanfi_veg_summary.R`, built by the gated
  `landweb.extended_analyses` target. The scan is multi-hour over the
  ~3.0 M km^2^ study area, so the result is cached as a `format = "file"` target
  and reused unless the CSV is deleted or the SCANFI/LTHFC version stamp changes.
