---
title: "LandWeb Study-Area Vegetation"
subtitle: "SCANFI species composition, and the kNN (v2) to SCANFI (v3) initial-condition change"
author: "Alex M. Chubaty"
date: today
format:
  pdf:
    pdf-engine: pdflatex
    toc: true
    toc-depth: 2
    number-sections: true
    fontsize: 11pt
fig-width: 8
fig-height: 6
bibliography:
  - ../citations/references.bib
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(targets)
library(data.table)
library(ggplot2)
library(kableExtra)

## display absolute pipeline paths as short project-relative paths
relpath <- function(p) sub("^.*/((?:inputs|outputs)/.*)$", "\\1", p, perl = TRUE)

## read from the pipeline (the literal tar_read() calls wire the tar_quarto deps);
## fall back to the seeded extended-analyses CSVs so the report also renders standalone.
scanfi_csv <- tryCatch(tar_read(scanfi_veg_summary_csv), error = function(e) {
  cand <- c("outputs/_extended_analyses/scanfi_lthfc_v10_2020_veg_summary.csv",
            "../outputs/_extended_analyses/scanfi_lthfc_v10_2020_veg_summary.csv")
  cand[file.exists(cand)][1]
})
knn_csv <- tryCatch(tar_read(knn_veg_summary_csv), error = function(e) {
  cand <- c("outputs/_extended_analyses/knn_lthfc_v10_2001_veg_summary.csv",
            "../outputs/_extended_analyses/knn_lthfc_v10_2001_veg_summary.csv")
  cand[file.exists(cand)][1]
})
veg <- fread(scanfi_csv)
knn <- fread(knn_csv)
area_cells <- veg$domain_forested_cells[1]
area_km2 <- veg$domain_area_km2[1]
suspects <- c("TSUG_HET", "THUJ_PLI", "PINU_CON_CON")
veg[, suspect := species %in% suspects]
pct <- function(code) veg[species == code, pct_forested]

## v2 (kNN) vs v3 (SCANFI): the LandWeb-species groups each product resolves
grp_scanfi <- veg[, .(SCANFI = paste(unique(common_name), collapse = ", ")), by = landweb_species]
grp_knn <- knn[, .(kNN = paste(unique(common_name), collapse = ", ")), by = landweb_species]
grp_cmp <- merge(grp_scanfi, grp_knn, by = "landweb_species", all = TRUE)
grp_cmp[is.na(kNN), kNN := "--- (not resolved)"]
setorder(grp_cmp, landweb_species)
scanfi_only <- setdiff(veg$landweb_species, knn$landweb_species)
n_grp_knn <- length(unique(knn$landweb_species))
n_grp_scanfi <- length(unique(veg$landweb_species))
```

\newpage

::: {.callout-warning}
**Draft report: pending human review**

Portions of this document were drafted with the help of generative AI (large language models) and have not yet been fully reviewed by the co-authors.
Content may be incomplete, inaccurate, or misattributed.
This notice will be removed once a complete review has been carried out.
:::

# Overview

This report summarizes the SCANFI v2 (2020) [@scanfi-v2] per-species
**crown-closure prevalence** over the **full LandWeb study area**
(`r format(round(area_km2), big.mark = ",")` km^2^,
`r format(area_cells, big.mark = ",")` forested 30 m cells). For each SCANFI
source species that maps to a simulated LandWeb species it reports the number
(and %) of forested study-area cells carrying *any* cover.

One primary purpose is to document the evidence behind the
**`THUJ_PLI`/`TSUG_HET` → `Abie_spp` merge** in `LandWeb_preamble`: western
redcedar and western hemlock are a *localized* SCANFI over-attribution in the
SW Alberta Rockies, not a study-area-wide signal.

It also documents the **change in starting conditions from v2 to v3**: the
underlying tree-species product moved from the Beaudoin *k*NN (2001) maps
[@BeaudoinEtAl2014] to SCANFI (2020) [@scanfi-v2] (see
[Initial conditions: Beaudoin kNN (2001) to SCANFI (2020)]).

# Study area composition

```{r}
#| label: tbl-composition
#| tbl-cap: "SCANFI species prevalence over the LandWeb study area, by cells with any cover (sorted). `LandWeb species` is the simulated species each SCANFI code merges into."
knitr::kable(
  veg[, .(
    `SCANFI code` = species,
    `Common name` = common_name,
    `LandWeb species` = landweb_species,
    `Cells (cover > 0)` = format(n_cover_gt0, big.mark = ","),
    `% of study area` = sprintf("%.3f", pct_forested)
  )],
  align = "lllrr",
  booktabs = TRUE
) |>
  kable_styling(font_size = 8, latex_options = "hold_position")
```

```{r}
#| label: fig-composition
#| fig-cap: "Prevalence (% of forested study-area cells with cover > 0) by SCANFI species; coastal-attribution suspects in red."
ggplot(veg, aes(reorder(common_name, pct_forested), pct_forested, fill = suspect)) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(values = c(`FALSE` = "grey40", `TRUE` = "firebrick"), guide = "none") +
  labs(x = NULL, y = "% of forested study-area cells with cover > 0") +
  theme_minimal(base_size = 12)
```

The composition is a textbook western-boreal signature --- black spruce
dominant, then tamarack, aspen, jack pine, and white spruce.

# Coastal over-attribution (the merge rationale)

```{r}
#| label: tbl-suspects
#| tbl-cap: "Coastal-attribution suspects: study-area prevalence and substantial-cover (>= 10%) extent."
knitr::kable(
  veg[species %in% suspects][
    order(-pct_forested),
    .(
      `SCANFI code` = species,
      `Common name` = common_name,
      `% study area (cover > 0)` = sprintf("%.3f", pct_forested),
      `% study area (>= 10%)` = sprintf("%.4f", pct_ge10),
      `Max cover %` = cover_max
    )
  ],
  align = "llrrr",
  booktabs = TRUE
) |>
  kable_styling(font_size = 8, latex_options = "hold_position")
```

Western hemlock (`TSUG_HET`) covers only **`r sprintf("%.2f", pct("TSUG_HET"))`%**
of the LandWeb study area, and western redcedar (`THUJ_PLI`)
**`r sprintf("%.2f", pct("THUJ_PLI"))`%** --- rounding errors at the study-area
scale. By contrast, a stand-alone count over the **Spray Lake FMA** (SW Alberta
Rockies) put western hemlock on **~25%** of pixels: a ~175× concentration in the
SW corner, exactly where SCANFI bleeds coastal species across the BC boundary.
Where these species do appear they can reach high cover (up to
`r veg[species == "TSUG_HET", cover_max]`% hemlock), but the footprint is tiny
and localized.

Because both fold into `Abie_spp` (the closest shade-tolerant softwood), the
merge:

- removes a stand-alone species that has no usable known-age / non-zero-biomass
  cohorts in the SW --- the cause of the `Biomass_borealDataPrep` non-conformable
  ("code 235") crash; and
- relabels only ~0.3% of the study area's cover, leaving the rest untouched.

`PINU_CON_CON` (shore pine, the coastal *Pinus contorta* variety) is effectively
absent across the study area
(`r format(veg[species == "PINU_CON_CON", n_cover_gt0], big.mark = ",")` cells),
so it is left merged into `Pinu_spp`, where --- like all pines --- it is harmless.

\newpage

# Initial conditions: Beaudoin kNN (2001) to SCANFI (2020)

The v2 model built its starting species/biomass/age conditions from the Beaudoin
*k*NN maps [@BeaudoinEtAl2014] --- 250 m, 2001 base year, derived by *k*-nearest-neighbours
from MODIS imagery and NFI photo-plots. v3 replaces these with SCANFI (2020)
[@scanfi-v2; @scanfi-methods] at 30 m. Because the model's initial vegetation is read
directly from this product, the switch changes the composition the simulation starts from,
in two ways: **which species are resolved**, and **their mapped extent**.

## Species the two products resolve

```{r}
#| label: tbl-resolved
#| tbl-cap: "Tree species each inventory resolves, grouped by simulated LandWeb species. SCANFI (v3) resolves larch and Douglas-fir --- absent from the kNN product (v2) --- and additionally splits Engelmann spruce and the two *Pinus contorta* varieties."
knitr::kable(
  grp_cmp[, .(
    `LandWeb species` = landweb_species,
    `kNN (v2, 2001)` = kNN,
    `SCANFI (v3, 2020)` = SCANFI
  )],
  align = "lll", booktabs = TRUE
) |>
  kable_styling(font_size = 8, latex_options = "hold_position") |>
  column_spec(2:3, width = "15em")
```

The kNN product resolves **`r n_grp_knn`** LandWeb species groups; SCANFI resolves
**`r n_grp_scanfi`**. The two SCANFI adds are consequential:
**`r paste(scanfi_only, collapse = " and ")`** --- i.e. **larch/tamarack** and
**Douglas-fir**, which have no kNN layer at all. SCANFI also resolves Engelmann spruce
(folded into `Pice_gla`) and both *Pinus contorta* varieties, where kNN carried a single
lumped fir group and a single lodgepole layer.

## Composition shift

```{r}
#| label: tbl-knn-composition
#| tbl-cap: "Beaudoin kNN (2001) per-species mean cover and prevalence over the LandWeb study area (sorted by mean cover)."
knitr::kable(
  knn[, .(
    `kNN species` = common_name,
    `LandWeb species` = landweb_species,
    `Mean cover %` = sprintf("%.2f", mean_cover),
    `% of domain (cover > 0)` = sprintf("%.2f", pct_domain)
  )],
  align = "llrr", booktabs = TRUE
) |>
  kable_styling(font_size = 8, latex_options = "hold_position")
```

```{r}
#| label: fig-knn-composition
#| fig-cap: "Beaudoin kNN (2001) mean cover by species over the LandWeb study area, coloured by simulated LandWeb species. The kNN product carries no larch or Douglas-fir layer."
ggplot(knn, aes(reorder(common_name, mean_cover), mean_cover, fill = landweb_species)) +
  geom_col() +
  coord_flip() +
  labs(x = NULL, y = "Mean cover (%) over the study area", fill = "LandWeb\nspecies") +
  theme_minimal(base_size = 12)
```

Both products agree on the dominant signature --- **black spruce** leads by a wide margin,
followed by aspen, jack pine, and white spruce. The most consequential difference is at #2:
in SCANFI, **tamarack/larch (`Lari_spp`) is the second most prevalent group** (on
**`r sprintf("%.1f", sum(veg[landweb_species == "Lari_spp", pct_forested]))`%** of the study
area), yet the kNN product has **no larch layer** --- so the v2 model started with that entire
LandWeb species group absent. The two products' resolutions (250 m vs 30 m) make the raw
prevalence values not directly comparable in magnitude, so the comparison here emphasizes the
species set and relative ordering rather than absolute percentages.

## Bearing on simulated outcomes

Because the initial cohort composition is seeded from this product, the v3 starting landscape
carries a **richer species pool** (`r n_grp_scanfi` vs `r n_grp_knn` LandWeb groups) with
larch and Douglas-fir newly present, Engelmann spruce distinguished, and a two-decade-newer
vintage (2001 to 2020) that already reflects the intervening disturbance and growth. These
initial-condition differences are one of the contributors to any change in simulated outcomes
between v2 and v3, alongside the fire-input update (report 02) and the code/module changes
(report 03); the output comparison (report 04) is where their combined effect is quantified.

::: {.callout-note}
**Scope of this comparison.** This section compares *species composition* only. Two related
initial-condition changes are out of scope here: (1) the **stand-age and biomass** layers
(kNN 250 m vs SCANFI 30 m), whose study-area-wide comparison needs the gated 30 m scan; and
(2) the **current-condition** update from the earlier CASFRI/Pickell-era layers to the 2026
merged Alberta AVI + BC VRI + national SBFI composite, which is pending that composite being
staged locally.
:::

\newpage

# References {.unnumbered}

::: {#refs}
:::

\newpage

# Provenance {.unnumbered}

- **Sources:** SCANFI v2 species crown-closure layers [@scanfi-v2], derived via
  the SCANFI approach [@scanfi-methods]
  (`SCANFI_spsCC_<code>_2020_v2_20260119.tif`); and the Beaudoin *k*NN 2001
  species layers [@BeaudoinEtAl2014]
  (`NFI_MODIS250m_2001_kNN_Species_<Sp>_v1.tif`). Both masked to the LandWeb
  study area (the LTHFC v10 union, `landweb_ltfc_v10`).
- **Data files:** `` `r relpath(scanfi_csv)` `` (SCANFI);
  `` `r relpath(knn_csv)` `` (kNN).
- **Method:** `R/scanfi_veg_summary.R` (multi-hour over the ~3.0 M km^2^ study
  area at 30 m) and `R/knn_veg_summary.R` (~40 s at 250 m), both cache-aware
  `format = "file"` targets built by the gated `landweb.extended_analyses`
  factory; reused unless the CSV is deleted or a version stamp in its name
  changes.

{{< include _provenance-appendix.qmd >}}
