---
title: "LandWeb Fire-Model Inputs"
subtitle: "The LandMine LTHFC (fire-return-interval) layer and its v10 update"
author: "Alex M. Chubaty"
date: today
format:
  pdf:
    pdf-engine: pdflatex
    toc: true
    toc-depth: 2
    number-sections: true
    fontsize: 11pt
fig-width: 8
fig-height: 6
bibliography:
  - ../citations/references.bib
csl: ../citations/ecology-letters.csl
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(targets)
library(data.table)
library(ggplot2)
library(kableExtra)

## display absolute pipeline paths as short project-relative paths
relpath <- function(p) sub("^.*/((?:inputs|outputs)/.*)$", "\\1", p, perl = TRUE)

## Per-run FRI attainment diagnostics, written by LandMine's summary events. Study-area
## specific, so they live under outputs/<studyArea>/mainSim/. Prefer the multi-replicate
## files; fall back to rep01. Missing artifacts degrade to a short note rather than an error.
.fri <- function(f) {
  roots <- c("outputs", "../outputs")
  cand <- unlist(lapply(roots, function(r) {
    d <- list.dirs(r, recursive = FALSE)
    c(file.path(d, "mainSim", f), file.path(d, "mainSim", "rep01", f))
  }))
  cand <- cand[file.exists(cand)]
  if (length(cand)) cand[1] else NA_character_
}
fri_diag_csv <- .fri("LandMine_FRI_diagnostics_multi.csv")
if (is.na(fri_diag_csv)) fri_diag_csv <- .fri("LandMine_FRI_diagnostics.csv")
fri_diag <- if (!is.na(fri_diag_csv)) fread(fri_diag_csv) else NULL

fri_map_png <- .fri("figures/LandMine/LandMine_FRI_zones_diagnostic_multi.png")
if (is.na(fri_map_png)) fri_map_png <- .fri("figures/LandMine/LandMine_FRI_zones_diagnostic.png")

## Read the LTHFC version summary from the pipeline; fall back to the seeded
## extended-analyses CSV so this report also renders standalone.
lthfc_csv <- tryCatch(
  tar_read(lthfc_version_summary_csv),
  error = function(e) {
    cand <- c(
      "outputs/_extended_analyses/lthfc_version_summary.csv",
      "../outputs/_extended_analyses/lthfc_version_summary.csv"
    )
    cand[file.exists(cand)][1]
  }
)
lthfc <- fread(lthfc_csv)

## coarse fire-return-interval bins (years); fri == 0 is the v8c non-burnable code
lthfc[, bin := cut(
  fri,
  breaks = c(-1, 0, 40, 60, 80, 120, Inf),
  labels = c("non-burnable", "1-40", "41-60", "61-80", "81-120", "121+")
)]

tot <- lthfc[, .(total_km2 = sum(area_km2)), by = version]
nonburn <- lthfc[fri == 0, .(nonburn_km2 = sum(area_km2)), by = version]
km <- function(v) format(round(tot[version == v, total_km2]), big.mark = ",")

## v8c -> v10 area by coarse bin (burnable only)
burn <- lthfc[fri > 0]
cmp <- dcast(burn[version %in% c("v8c", "v10")], bin ~ version,
  value.var = "area_km2", fun.aggregate = sum
)
## dcast orders columns alphabetically (v10 before v8c); force v8c, v10 so the
## positional col.names in tbl-area-bin line up with the data.
setcolorder(cmp, c("bin", "v8c", "v10"))
band <- function(b, v) cmp[bin == b][[v]]
```

\newpage

::: {.callout-warning}
**Draft report: pending human review**

Portions of this document were drafted with the help of generative AI (large language models) and
have not yet been fully reviewed by the co-authors.
Content may be incomplete, inaccurate, or misattributed.
This notice will be removed once a complete review has been carried out.
:::

# Overview

LandWeb simulates wildfire with an implementation of **LandMine** [@Andison:1996; @Andison:1998].
The single most important spatial input to that fire model is the **LTHFC** (Long-Term Historic
Fire Cycle) polygon layer: each polygon carries an integer **fire-return interval (FRI, years)**
that sets how often that part of the landscape burns. This report documents the current **v10**
LTHFC layer --- in particular the **June 2026 Northwest Alberta (NWAB) update**
[@peterson-nwab-data; @peterson-nwab-final] --- and compares it against the previously used **v8c**
layer and the **Spray Lake Sawmills (SLS v3)** variant.

The key message is a clean modelling seam: **v10 changes only the fire-return interval assigned to
each zone. Everything else in the fire model is unchanged.** Any change in simulated fire regime
attributable to inputs therefore traces to the FRI-by-zone values documented here.

# How LandMine uses the LTHFC layer

The LTHFC polygons enter the pipeline as a fire-return-interval raster and drive ignition
frequency:

- `LandWeb_preamble` downloads the v10 layer, renames its FRI field to `fireReturnInterval`, cleans
  the polygons (`LandWebUtils::polygonClean(type = "LandWeb", minFRI = ...)`), and rasterizes
  `fireReturnInterval` onto the study-area grid.
- `LandMine` consumes `fireReturnInterval` (years) together with a separate flammability mask
  (`rstFlammable`). The number of fires ignited per year in a zone is derived from that zone's area
  and its FRI (zone area / mean fire size / FRI).

Crucially, **only the fire-return interval comes from the LTHFC layer.** The fire-size distribution
(a Truncated Pareto), the rate of spread (`ROSTable`, `ROSother`), and the per-pixel spread
probability all come from LandMine's own parameters and its `DEoptim` calibration table --- not from
the LTHFC polygons. This is why a v8c-to-v10 swap is cleanly interpretable: it moves area between
FRI classes and nothing else.

# Fire-spread calibration

The spread parameters named above --- the four `spawnNewActive` probabilities, the two
`sizeCutoffs`, and `spreadProb` --- are not measured. They are fitted by `DEoptim` against the
landscape statistics @Andison:1996 reported, on a uniform, fully-flammable synthetic landscape so
the fit is not conditioned on any one study area's fuel pattern.

Two properties of that fit matter for reading everything below.

**The targets are defined in hectares, and Andison rasterized at 50 m (quarter-hectare) pixels.**
LandWeb runs much coarser, so fire sizes are specified in hectares and converted per resolution
(`LandWebUtils::landmine_optim_fireSizes()`). A fixed pixel-count vector would mean a different
physical fire at each resolution, which is what previously placed the 240 m calibration outside
the 10--3,000 ha range the shape relationship was fitted over.

**Remnant islands cannot be reproduced equally at all resolutions.** Andison's Table 3.6 puts 65 %
of island area in the 2--5 ha class for fires under 1,000 ha; at 240 m a single pixel is already
5.76 ha, so that class cannot exist and the island target is unreachable however the parameters are
set. The island term is therefore down-weighted with coarsening resolution
(`LandWebUtils::landmine_optim_islandWeight()`, `min(1, 50 / pixelSize)`), so the optimizer is not
driven by a statistic the grid forbids at the expense of ones it can reproduce.

```{r}
#| label: calib-load
.calibDir <- c("outputs/_calibration", "../outputs/_calibration")
.calibDir <- .calibDir[dir.exists(.calibDir)][1]
.readCalib <- function(res, what) {
  if (is.na(.calibDir)) return(NULL)
  f <- file.path(.calibDir, sprintf("landmine_%s_%dm.rds", what, res))
  if (file.exists(f)) readRDS(f) else NULL
}
.calibRes <- c(240, 120)
.deopt <- stats::setNames(lapply(.calibRes, .readCalib, what = "DEoptim_result"), .calibRes)
.comps <- stats::setNames(lapply(.calibRes, .readCalib, what = "calib_components"), .calibRes)
.haveCalib <- any(!vapply(.deopt, is.null, logical(1)))
```

```{r}
#| label: calib-guard
#| results: asis
if (!.haveCalib) {
  cat(
    "::: {.callout-warning}\n",
    "## Calibration outputs not present\n\n",
    "No `landmine_DEoptim_result_*.rds` was found under `outputs/_calibration/`. ",
    "Run the calibration (`LandWebUtils::landmine_optim_calibrate()`) to populate this section; ",
    "the figures below are regenerated from those artifacts on every render.\n",
    ":::\n",
    sep = ""
  )
}
```

## Calibrated parameters

```{r}
#| label: tbl-calib-params
#| tbl-cap: "Fitted fire-spread parameters by resolution. `spawnNewActive` and `sizeCutoffs` are fitted as base-10 logarithms and shown here on the natural scale. Fire sizes were held identical in hectares across resolutions, so differences reflect resolution, sampling noise, or both --- see the caveats below."
if (.haveCalib) {
  .rows <- lapply(names(.deopt), function(r) {
    d <- .deopt[[r]]
    if (is.null(d)) return(NULL)
    p <- LandWebUtils::landmine_optim_unpack(d$optim$bestmem)
    data.frame(
      resolution = paste0(r, " m"),
      spawn1 = p$spawnNewActive[1], spawn2 = p$spawnNewActive[2],
      spawn3 = p$spawnNewActive[3], spawn4 = p$spawnNewActive[4],
      cutoff1_ha = p$sizeCutoffs[1], cutoff2_ha = p$sizeCutoffs[2],
      spreadProb = p$spreadProb,
      islandWeight = LandWebUtils::landmine_optim_islandWeight(as.numeric(r)),
      objective = d$optim$bestval
    )
  })
  .tbl <- do.call(rbind, .rows)
  knitr::kable(.tbl, digits = c(0, 4, 4, 4, 4, 0, 0, 3, 3, 4))
}
```

## Convergence

```{r}
#| label: fig-calib-convergence
#| fig-cap: "Best-of-population objective across DEoptim generations, with a dotted line at the last improvement. A flat tail means the run converged; a still-falling curve means it hit `itermax` while improving. On these fits improvement stops well before the end, so iteration count is not the limiting factor --- objective noise is, which is what replication addresses."
#| fig-width: 7
#| fig-height: 3.5
if (.haveCalib) {
  for (r in names(.deopt)) {
    if (is.null(.deopt[[r]])) next
    print(
      LandWebUtils::landmine_plot_calibConvergence(.deopt[[r]]) +
        ggplot2::labs(title = sprintf("Calibration convergence (%s m pixels)", r))
    )
  }
}
```

## Fit against Andison's statistics

```{r}
#| label: fig-calib-shape
#| fig-cap: "Simulated fire shape against the relationship Andison fitted, `SHAPE = 1.770 + 0.041 log10(AREA)^4`. The solid line is his **empirical** fit; the dashed line is the fit to his own model's output, shown only for context --- targeting it would be circular. The shaded band marks the 10--3,000 ha range the relationship was fitted over. Small fires at coarse resolution are quantization-limited: a 50 ha fire is only 9 pixels at 240 m and cannot express a convoluted perimeter."
#| fig-width: 7
#| fig-height: 4.5
if (.haveCalib) {
  for (r in names(.comps)) {
    if (is.null(.comps[[r]])) next
    print(
      LandWebUtils::landmine_plot_calibShape(.comps[[r]]) +
        ggplot2::labs(title = sprintf("Fire shape vs. area (%s m pixels)", r))
    )
  }
}
```

```{r}
#| label: fig-calib-islands
#| fig-cap: "Simulated remnant-island fraction against Andison Table 3.5 (empirical column; the model column of that table is not a target). Islands under 2 ha are excluded, following Andison. This statistic is resolution-limited by construction --- see the down-weighting note above --- so departures at coarse resolution are expected rather than diagnostic of a poor fit."
#| fig-width: 7
#| fig-height: 4.5
if (.haveCalib) {
  for (r in names(.comps)) {
    if (is.null(.comps[[r]])) next
    print(
      LandWebUtils::landmine_plot_calibIslands(.comps[[r]]) +
        ggplot2::labs(title = sprintf("Remnant islands (%s m pixels)", r))
    )
  }
}
```

## How much to trust these values

Three caveats belong with any use of the fitted parameters.

**The objective identifies some parameters and not others.** Fitted against the previous objective,
43 repeat runs of the same configuration scattered four of the seven parameters across more than
90 % of their search interval, and repeat solutions proved statistically indistinguishable from one
another --- and from some randomly drawn parameter vectors. The current objective adds constraints
and does separate fitted from random parameters, but it still cannot rank near-equivalent solutions.
Treat any single fit as one draw from a broad optimum rather than as a point estimate.

**A stochastic objective invites over-fitting to its own random draws.** Common random numbers make
parameter vectors comparable, but let the optimizer exploit the particular realization it was given.
Re-evaluating a fit at independent seeds is the check that matters, and is reported with each run.

**Fire sizes here are a heuristic, not a fit to observed fires.** The truncated-Pareto shape
governing fire sizes is set by a rule of thumb ("95 % of area in 5 % of fires"), not fitted to NBAC
or NFDB. Those records are used to *compare against* simulated output in the burn summaries, but the
simulated size distribution was never fitted to them --- so agreement there is a weaker check than it
may appear.

# The v10 update: Northwest Alberta NRV (2026)

The v10 layer incorporates a major revision to the fire regime of **northwest Alberta**, produced by
braiding Western science with Indigenous and long-time-resident knowledge
[@peterson-nwab-data; @peterson-nwab-final]. It replaces the original characterization from the
LandWeb pre-industrial fire-regime synthesis [@Andison:2019], which had assigned NW Alberta a
**default 100-year** fire cycle --- a value the synthesis itself flagged as lowest-confidence and
likely too long.

## Method (in brief)

1. **Time-since-fire** reconstruction from four stand-origin layers (reference years 1990 and 2023,
   each in *Liberal* and *Conservative* confidence in mapped fires), built from a hierarchical
   source-priority rule set (fire-history records, then undisturbed AVI, pre-harvest AVI, minimum
   harvest age, adjacency, and the 1949--1956 Broad-Scale Inventory).
2. **Fire cycle** estimated as the slope of cumulative undisturbed area against time-since-fire on a
   semi-log plot [the method of Reed et al. 1998, per @peterson-nwab-data]. An inflection near
   **1960** separates the pre-industrial regime (fire cycles of about 43--53 yr) from the
   fire-control era (about 354--460 yr).
3. **Three scenarios** bracket the estimate --- *Shortest*, *Intermediate*, and *Longest* --- cut to
   Canadian eco-districts so the zones integrate seamlessly into the larger LandWeb map.

The project recommends, and LandWeb adopts, the **Intermediate** scenario.

## Fire-return intervals by natural subregion

```{r}
#| label: tbl-fri-nsr
#| tbl-cap: "NWAB fire-return intervals (years) by Alberta natural subregion, per scenario. Reproduced from @peterson-nwab-data, Table 2-5; the adopted Intermediate column is in bold."
fri_nsr <- data.frame(
  NSR = c("Boreal Subarctic", "Central Mixedwood", "Dry Mixedwood",
          "Lower Boreal Highlands", "Northern Mixedwood", "Upper Boreal Highlands"),
  Shortest = c(45, 45, 35, 45, 45, 35),
  Intermediate = c(55, 55, 40, 55, 55, 45),
  Longest = c(65, 65, 45, 65, 65, 55)
)
knitr::kable(fri_nsr, align = "lrrr", booktabs = TRUE,
  col.names = c("Natural subregion", "Shortest", "Intermediate (adopted)", "Longest")) |>
  kable_styling(font_size = 9, latex_options = "hold_position") |>
  column_spec(3, bold = TRUE)
```

Against the original 100-year default, the adopted intervals of **40--55 years** roughly *halve*
the fire cycle across NW Alberta, which (per the Final Report) raises the modelled proportion of
young forest and lowers old forest in the natural range of variation for that region.

# Version comparison: v10 vs v8c vs SLS v3

This section is derived directly from the LTHFC shapefiles in `inputs/` by
`R/lthfc_summary.R` (area-by-FRI over each layer). The three layers span slightly different extents
and schemas: **v10** (`r km("v10")` km^2^, field `LTFC10`, LCC projection), **v8c**
(`r km("v8c")` km^2^, field `LTHFC`, UTM 11N), and **SLS v3**
(`r km("sls_v3")` km^2^, field `LTHFC`, UTM 11N).

## Where the change is

```{r}
#| label: fig-lthfc-map
#| fig-cap: "Fire-return interval across the LandWeb study area under v8c (previous) and v10 (2026). Colour runs red (short fire cycle) to yellow (long), following the NWAB mapping convention. The Northwest Alberta area revised in 2026 is outlined in black: inside it, v10 shifts toward red (shorter intervals), while the rest of the map is largely unchanged."
#| fig-width: 8
#| fig-height: 5.5
library(sf)
gpkg <- tryCatch(
  tar_read(lthfc_change_map_gpkg),
  error = function(e) {
    cand <- c(
      "outputs/_extended_analyses/lthfc_change_map.gpkg",
      "../outputs/_extended_analyses/lthfc_change_map.gpkg"
    )
    cand[file.exists(cand)][1]
  }
)
mp <- st_read(gpkg, layer = "lthfc", quiet = TRUE)
mp$band <- cut(mp$fri, breaks = c(0, 40, 60, 80, 120, Inf),
  labels = c("1-40", "41-60", "61-80", "81-120", "121+"))
mp$version <- factor(mp$version, levels = c("v8c (previous)", "v10 (2026)"))
aoi <- tryCatch(st_read(gpkg, layer = "aoi", quiet = TRUE), error = function(e) NULL)
prov <- tryCatch(st_read(gpkg, layer = "provinces", quiet = TRUE), error = function(e) NULL)

## red (short cycle) -> yellow (long cycle), per the NWAB convention
band_cols <- setNames(
  colorRampPalette(c("red", "yellow"))(5),
  c("1-40", "41-60", "61-80", "81-120", "121+")
)

p <- ggplot()
if (!is.null(prov)) p <- p + geom_sf(data = prov, fill = "grey96", color = "grey75", linewidth = 0.2)
p <- p + geom_sf(data = mp, aes(fill = band), color = NA)
if (!is.null(prov)) p <- p + geom_sf(data = prov, fill = NA, color = "grey40", linewidth = 0.2)
if (!is.null(aoi)) p <- p + geom_sf(data = aoi, fill = NA, color = "black", linewidth = 0.7)
bb <- sf::st_bbox(mp)
p +
  scale_fill_manual(values = band_cols, name = "Fire-return\ninterval (yr)", drop = FALSE) +
  coord_sf(xlim = bb[c("xmin", "xmax")], ylim = bb[c("ymin", "ymax")]) +
  facet_wrap(~version) +
  theme_bw(base_size = 11) +
  theme(axis.text = element_text(size = 6), legend.position = "right") +
  labs(x = NULL, y = NULL)
```

The two versions are near-identical outside Northwest Alberta; within the outlined AOI, v10 shifts
toward red (shorter fire cycles). The difference map below isolates that change directly.

```{r}
#| label: fig-lthfc-diff
#| fig-cap: "Change in fire-return interval, v10 minus v8c (years), over areas burnable in both versions. Red = v10 shortened the fire cycle (more fire); blue = lengthened; grey = unchanged. The change is concentrated in the outlined Northwest Alberta AOI, where the 2026 update roughly halved the cycle."
#| fig-width: 7
#| fig-height: 5.5
library(terra)
library(tidyterra)
dtif <- tryCatch(
  tar_read(lthfc_diff_tif),
  error = function(e) {
    cand <- c(
      "outputs/_extended_analyses/lthfc_diff.tif",
      "../outputs/_extended_analyses/lthfc_diff.tif"
    )
    cand[file.exists(cand)][1]
  }
)
dr <- terra::rast(dtif)

pd <- ggplot()
if (!is.null(prov)) pd <- pd + geom_sf(data = prov, fill = "grey96", color = "grey75", linewidth = 0.2)
pd <- pd + tidyterra::geom_spatraster(data = dr, maxcell = 5e5)
if (!is.null(prov)) pd <- pd + geom_sf(data = prov, fill = NA, color = "grey40", linewidth = 0.2)
if (!is.null(aoi)) pd <- pd + geom_sf(data = aoi, fill = NA, color = "black", linewidth = 0.7)
pd +
  scale_fill_gradient2(
    low = "red", mid = "grey92", high = "blue", midpoint = 0,
    na.value = "transparent", name = "FRI change\n(yr; v10 - v8c)"
  ) +
  coord_sf(xlim = bb[c("xmin", "xmax")], ylim = bb[c("ymin", "ymax")]) +
  theme_bw(base_size = 11) +
  theme(axis.text = element_text(size = 6)) +
  labs(x = NULL, y = NULL)
```

## Area by fire-return interval

```{r}
#| label: tbl-area-bin
#| tbl-cap: "Study-area area (thousand km2) by fire-return-interval band (years) under v8c vs v10 (burnable polygons only)."
cmp_tbl <- copy(cmp)
for (v in c("v8c", "v10")) set(cmp_tbl, j = v, value = round(cmp_tbl[[v]] / 1000, 1))
knitr::kable(cmp_tbl, align = "lrr", booktabs = TRUE,
  col.names = c("FRI band (yr)", "v8c (1000 km2)", "v10 (1000 km2)")) |>
  kable_styling(font_size = 9, latex_options = "hold_position")
```

```{r}
#| label: fig-area-bin
#| fig-cap: "Study-area area by fire-return-interval band, v8c vs v10 (burnable polygons only). v10 shifts area out of the longest intervals and into the 41--60 yr band --- the NW Alberta halving."
pd <- melt(cmp, id.vars = "bin", variable.name = "version", value.name = "area_km2")
pd <- pd[!is.na(bin)]
ggplot(pd, aes(bin, area_km2 / 1000, fill = version)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.75) +
  scale_fill_manual(values = c(v8c = "grey60", v10 = "firebrick"),
    labels = c(v8c = "v8c (previous)", v10 = "v10 (2026)")) +
  labs(x = "Fire-return interval (years)", y = "Area (thousand km2)", fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top")
```

The signature of the NW Alberta update is visible at the landscape scale: relative to v8c, v10 moves
roughly **`r round((band("41-60","v10") - band("41-60","v8c"))/1000)` thousand km^2^** of
additional area into the **41-60 year** band (`r round(band("41-60","v8c")/1000)` to
`r round(band("41-60","v10")/1000)` thousand km^2^), drawing it down from the **81-120**
and **over-120** year bands. Note this is a *partial* update: v10 revises NW Alberta only, so most of
the map retains its earlier intervals, and full re-integration of the Intermediate scenario across
the whole LandWeb LTHFC map remains outstanding [@peterson-nwab-final].

## The schema and non-burnable-coding change

The v10 layer is not just new values --- it is a new schema. The FRI field was renamed
(`LTHFC` to `LTFC10`), the projection changed (UTM 11N to LCC), and non-burnable land is
handled differently: **v8c encoded `r format(round(nonburn[version == "v8c", nonburn_km2]), big.mark = ",")` km^2^
as fire-return interval `0`**, whereas v10 carries no `0` class and instead marks water via a lakes
attribute. Downstream, `minFRI` in the preamble still maps very-short intervals to non-burnable; with
v10's 35-year minimum nothing is dropped at the pipeline's current `minFRI = 25`.

The **SLS v3** layer is the Spray Lake Sawmills variant of the map (a near-full-extent layer, not a
local patch); it was the source from which the Spray Lake / Crowsnest polygons were taken to build
the interim `v8d` patch (`R/new_lthfc.R`). It is included here for the requested comparison but is
not the current pipeline input.

\newpage

# Fire-return-interval attainment

Whether the simulation actually *delivers* the intervals mapped above is a separate question from
what the layer says. LandMine writes a per-zone attainment check every run; this section reads it.

```{r}
#| label: fri-verdict
#| results: asis
if (is.null(fri_diag)) {
  cat("_No FRI diagnostics found. They are written by LandMine's summary events; ",
      "run the main simulation to populate them._\n", sep = "")
} else {
  cat("Source: `", relpath(fri_diag_csv), "`.\n\n", sep = "")
  cat(LandWebUtils::landmine_fri_verdict(fri_diag), "\n", sep = "")
}
```

```{r}
#| label: tbl-fri-attainment
#| tbl-cap: "Achieved versus target fire-return interval by zone. `ratio` is achieved/target over the area the fire model can actually burn, so a value above 1 means the zone burns too rarely. `% in study area` is the share of each zone inside that burnable area: where it is low, the uncorrected ratio (`ratio, all pixels`) is badly inflated, because pixels the model cannot reach still sit in the denominator."
if (!is.null(fri_diag)) {
  tab <- fri_diag[, .(
    `target (yr)` = LTHFC,
    `achieved (yr)` = round(FRI, 1),
    ratio = round(ratio, 2),
    status,
    `% in study area` = round(pctInStudyArea, 1),
    `% flammable` = round(pctFlam, 1),
    `ratio, all pixels` = round(ratioUnmasked, 2)
  )]
  kbl(tab, booktabs = TRUE, linesep = "") |>
    kable_styling(latex_options = c("hold_position", "scale_down"))
}
```

```{r}
#| label: fig-fri-drivers
#| fig-cap: "What separates any off-target zones from the rest. Points are FRI zones, labelled by target interval; above the dotted line a zone burns too rarely. The y axis is log2, so equal distances are equal factors."
#| fig-width: 8
#| fig-height: 4
if (!is.null(fri_diag)) {
  LandWebUtils::landmine_plot_fri_drivers(fri_diag, "")
}
```

```{r}
#| label: fig-fri-zones
#| fig-cap: "FRI zones by target interval (left) and by achieved/target ratio (right). Red marks zones burning too rarely, blue too often; near-neutral is on target."
if (!is.na(fri_map_png)) {
  knitr::include_graphics(normalizePath(fri_map_png))
}
```

::: {.callout-note}
## Reading the two ratio columns

LandMine masks both its rate-of-spread map and its spread probability to the study area, so a
pixel outside that polygon can be neither ignited nor spread into. The fire-return-interval raster
is not similarly masked and routinely overhangs the polygon. Any such pixel therefore enters a
zone's achieved-interval denominator while contributing no opportunity to burn, which is what
`ratio, all pixels` reports and what earlier `LandMine_FRI_summary.csv` files contain. Where
`% in study area` is well below 100, the two columns diverge sharply, and only `ratio` is
meaningful.
:::

# Expected effect on the simulated fire regime

Because LandMine derives ignition frequency directly from FRI, shortening NW Alberta's fire cycle
increases the number of fires there, which over a multi-century run raises the proportion of
young/recently-burned forest and lowers old forest --- shifting the seral-stage natural range of
variation for that region. Quantifying that shift in simulated outputs (v2 vs v3) is the subject of
the output-comparison report.

\newpage

# References {.unnumbered}

::: {#refs}
:::

\newpage

# Provenance {.unnumbered}

- **v10 source:** the NWAB Intermediate fire-return intervals [@peterson-nwab-data], cut into the
  LandWeb LTHFC map; `inputs/landweb_ltfc_v10.shp` (field `LTFC10`). Original fire-cycle
  characterization from @Andison:2019.
- **Comparison layers:** `inputs/landweb_ltfc_v8c.shp` (field `LTHFC`),
  `inputs/ltfc_sls_v3.shp` (field `LTHFC`).
- **Data file:** `` `r relpath(lthfc_csv)` ``
- **Method:** `R/lthfc_summary.R`, built by the gated `lthfc_version_summary_csv` target. Areas are
  geodesic (each layer reprojected to EPSG:4326, then `sf::st_area`) so the differently-projected
  layers are comparable. The scan is light and cached as a `format = "file"` target; delete the CSV
  to recompute.
- **FRI-by-NSR table** reproduced from @peterson-nwab-data, Table 2-5.

{{< include _provenance-appendix.qmd >}}
