---
title: "LandWeb v3: Provenance"
subtitle: "Input-data manifest, bibliography sync, and reproducibility (design in progress)"
author: "Alex M. Chubaty"
date: today
format:
  pdf:
    pdf-engine: pdflatex
    toc: true
    toc-depth: 2
    number-sections: true
    fontsize: 11pt
bibliography:
  - ../citations/references.bib
csl: ../citations/ecology-letters.csl
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(data.table)
library(kableExtra)
nn <- function(x) if (is.null(x) || !length(x)) "" else as.character(x)

mf <- c("_input_manifest.json", "../_input_manifest.json")
mf <- mf[file.exists(mf)][1]
j <- jsonlite::read_json(mf)
rec <- rbindlist(lapply(j$records, function(r) {
  data.table(
    id = r$id, name = r$name, source = nn(r$source$type),
    version = nn(r$version_or_vintage), license = nn(r$license)
  )
}))
setorder(rec, id)

## caribou reference artifacts (study-area-independent -> outputs/_reference/); regenerate with
## scripts/make_caribou_reference.R. Paths tried from both the project root and reports/.
.ref <- function(f) {
  cand <- file.path(c("outputs/_reference", "../outputs/_reference"), f)
  cand <- cand[file.exists(cand)]
  if (length(cand)) cand[[1L]] else NA_character_
}
cb_png <- .ref("caribouRanges_report.png")
cb_csv <- .ref("caribouRanges.csv")
cb_tab <- if (!is.na(cb_csv)) fread(cb_csv) else NULL
if (!is.null(cb_tab)) setorder(cb_tab, layer, source)
cb_eccc <- if (!is.null(cb_tab)) cb_tab[layer == "eccc"] else NULL
## The regional total is a row the script emits, not the column sum: a cross-border range appears
## under two jurisdictions but is ONE reporting unit, since the summaries group by name.
cb_tot <- if (!is.null(cb_tab)) cb_tab[layer == "regional" & source == "(all sources)"] else NULL
if (!is.null(cb_tot) && !nrow(cb_tot)) cb_tot <- NULL
```

\newpage

::: {.callout-warning}
**Draft report: pending human review**

Portions of this document were drafted with the help of generative AI and have not been fully
reviewed. The input manifest, the bibliography sync, and the per-report build-provenance appendix
are **implemented**; the **output manifest** and the **ODD / TRACE narratives** are designed here
but not yet fully built (see the relevant sections).
:::

# Overview

This report documents how LandWeb v3 records the **provenance of both its inputs and its outputs** ---
where each dataset came from, and how each result can be traced back to the data + code + environment
that produced it. **Inputs and outputs are documented together here** (rather than in separate
reports): they are the two halves of one end-to-end lineage, and a reader auditing a result needs
both in one place. The design has five parts, at different stages of maturity:

1. an **input-data manifest** (implemented) --- one machine-readable record per input dataset (below);
2. an **auto-synced bibliography** (implemented) --- data-source citations derived from the manifest;
3. a **per-report build-provenance appendix** (implemented) --- repo commit, pinned module/package
   revisions, and software environment stamped into every rendered report (see the appendix to this
   document, and the shared `_provenance-appendix.qmd` include);
4. **output provenance and lineage** (design below; partly implemented) --- each output artifact tied
   to its producing `targets` target, the module revisions that generated it, its input lineage, and
   the run parameters (study area, replicates, seed); and
5. **reproducibility-standards alignment** (below) --- how the above maps onto FAIR, PERFICT, the Ten
   Simple Rules, ODD, and TRACE.

# Input-data manifest

The manifest (`` `r mf` ``) is a schema.org `DataCatalog` (JSON-LD) with one record per input
dataset, built with the `workflowtools` manifest API (`input_manifest_record()` /
`register_input()`) from `R/build_input_manifest.R`. Each record carries a stable `id`, the source
(type + URL), a `retrieved_at` timestamp, the local path, version/vintage, license, and a citation
pointer. It currently registers **`r nrow(rec)`** datasets:

```{r}
#| label: tbl-manifest
#| tbl-cap: "Registered input datasets (from _input_manifest.json). Each record additionally stores its retrieved_at timestamp, local path, license, and citation pointer."
knitr::kable(
  rec[, .(ID = id, Dataset = name, Source = source, `Version / vintage` = version)],
  align = "llll", booktabs = TRUE
) |>
  kable_styling(font_size = 8, latex_options = "hold_position") |>
  column_spec(2, width = "20em") |>
  column_spec(4, width = "12em")
```

The manifest is hand-curated (run `R/build_input_manifest.R` to regenerate) and committed to the
repo, so it travels with the code. It deliberately omits content hashes for the multi-GB rasters;
`sha256` can be added per record where a hash is worth carrying.

One group of records is **derived rather than hand-listed**: the six caribou range sources are
generated from `LandWebUtils::caribouRangeLayers()`. That is the deliberate exception to the rule
above, for the reason given in the next section --- they are the inputs most likely to change, and a
hand-copied list would drift from the code that actually fetches them.

# Multi-source assembly: caribou ranges {#sec-caribou-provenance}

Most inputs are one dataset from one authority. The caribou reporting ranges are the exception, and
they are worth documenting in full because they exercise every part of the provenance problem at once:
**six** sources, six authorities, six vintages, six licences (one of them not openly licensed at all),
disagreeing attribute schemas, disagreeing spellings for the same feature, and a national alternative
that could plausibly have replaced the lot. They are also the input most likely to change --- any of the
six jurisdictions can re-delineate independently --- so how the assembly is recorded matters more here
than anywhere else.

## Why assembled here rather than delivered pre-combined

Earlier LandWeb versions consumed a **hand-combined** caribou layer. Auditing it against its own
documented sources found that the manual combine had:

* **dropped the Northwest Territories ranges entirely** --- 0 of the delivered layer's 74 features intersect either NWT tenure, so the NWT tenure × caribou reporting units could not be rebuilt at all;
* left range names scattered across **five** per-jurisdiction columns (`RANGE_NAME` populated for only the Ontario and Manitoba features, British Columbia in `HERD_NAME`, Alberta in `LOCALRANGE`, Saskatchewan in `CONUNIT`/`RNGEUNIT`), so labelling on any single column silently discarded most of the ranges; and
* mis-cited one source --- the Saskatchewan URL of record pointed at a raster habitat-potential product rather than the vector ranges layer actually delivered.

None of these are visible in a combined deliverable; all three are structural consequences of
combining by hand. Assembling from the sources in-pipeline
(`LandWebUtils::buildCaribouRanges()`) removes the whole class: each source is fetched, filtered and
labelled on its own terms, and a bad merge becomes a one-row fix rather than a re-delivery.

The assembly also applies two corrections that belong in code, not in a spreadsheet:

* **locally extirpated herds are excluded.** Two sources flag this, in different vocabulary --- Alberta `STATUS == "Extirp"` (the Banff range, lost to an avalanche in 2009) and British Columbia `HERD_STATUS == "Extirpated"`. British Columbia's field is `HERD_STATUS` over its live WFS but truncated to `HERD_STAT` in a shapefile export, so both spellings are candidates.
* **range names are corrected where a source is demonstrably wrong** (`Bischto` → `Bistcho`). This is not cosmetic: the summaries group features by **name**, so a range two provinces spell differently is reported as two units --- and it is the same mechanism that stitches three cross-border ranges (`Chinchaga`, `Narraway`, `Redrock-Prairie Creek`) back into single reporting units for free. The bar is deliberately **evidence of an error, not disagreement between sources**, since otherwise this step would quietly overwrite one jurisdiction's nomenclature with a neighbour's. Each correction is checked against the authoritative toponym in NRCan's Canadian Geographical Names Database: it lists **Bistcho Lake** and no place called "Bischto". Two candidates were rejected on the same test --- British Columbia's `Snake-Sahtaneh` is *correct* (CGNDB has **Sahtaneh River**, and no "Sahtahneh"), so the federal layer carries that typo, not BC; and Alberta's `East Side Athabasca` / `West Side Athabasca` are **official enumerated domain values** in its published metadata rather than truncations of ECCC's `... River` forms.

## The national alternative, and why it is not used

Environment and Climate Change Canada publishes a national caribou layer that could in principle
replace all six. @fig-caribou-sources compares them. With designatable-unit scope matched (boreal and
mountain caribou; barren-ground and Peary are not used in LandWeb), the two **cover essentially the
same ground** --- the footprints agree to within about 2% --- but partition it differently: the
assembled layer resolves `r if (!is.null(cb_tot)) cb_tot$n_units else 69` reporting units against the
federal layer's `r if (!is.null(cb_eccc)) cb_eccc$n_units else 52`.

So the choice is not about coverage. It is about **granularity and vocabulary**, and both point the
same way: LandWeb reports to *jurisdictional* partners, each of whom manages and names their own
herds. The federal Southern Mountain units are aggregate Local Population Units that **merge** the
finer provincial herds, and for ranges lying wholly inside one province the provincial polygon is
geometrically identical to the federal one anyway (Alberta scores IoU = 1.0000 against it for 9 of 12
boreal ranges). Adopting the national layer would therefore hand partners coarser units under names
they do not use, in exchange for no additional information.

What is given up is real and worth recording: the federal layer carries whole **cross-border** ranges
as single units, where each provincial source holds only its own share, and it is the join key to
ECCC's anthropogenic disturbance-footprint products. It is retained as a documented comparison --- the
right-hand panel below, regenerated by `scripts/make_caribou_reference.R` --- rather than as a
reporting layer.

```{r}
#| label: fig-caribou-sources
#| fig-cap: "Caribou reporting ranges: the six-source jurisdictional assembly LandWeb reports on (left) against ECCC's national Priority Species at Risk layer (right), both clipped to the LandWeb v10 area of interest and restricted to boreal and mountain caribou with locally extirpated herds removed. Fill identifies the contributing jurisdiction; hatching marks mountain caribou. The two layers cover nearly the same ground but partition it differently --- note the federal panel's coarser mountain units, which aggregate the finer provincial herds. The left panel is built by the same `LandWebUtils::buildCaribouRanges()` the pipeline uses, so the figure cannot drift from the reporting units."
#| out-width: "100%"
## Pre-rendered by scripts/make_caribou_reference.R rather than rebuilt here: it fetches six remote
## sources (one of them Drive-authenticated) and would otherwise duplicate the plotting code.
if (!is.na(cb_png)) {
  knitr::include_graphics(normalizePath(cb_png))
} else {
  knitr::asis_output("*(caribou comparison map not found --- run `scripts/make_caribou_reference.R`)*")
}
```

```{r}
#| label: tbl-caribou-sources
#| tbl-cap: "Caribou range units by contributing source, within the LandWeb v10 area of interest. Only Alberta and British Columbia contribute mountain caribou; the remaining jurisdictions are boreal-only here. The `eccc` row is the national comparison layer, not a reporting input."
if (!is.null(cb_tab)) {
  knitr::kable(
    cb_tab[, .(Layer = layer, Source = source, Units = n_units,
               Boreal = n_boreal, Mountain = n_mountain,
               `Area (km^2^)` = format(area_km2, big.mark = ","))],
    align = "llrrrr", booktabs = TRUE
  ) |>
    kable_styling(font_size = 9, latex_options = "hold_position")
} else {
  knitr::asis_output("*(caribou source table not found --- run `scripts/make_caribou_reference.R`)*")
}
```

Per-source provenance --- endpoint, label column, status field and extirpation rule --- is documented
on `?caribouRangeLayers` and registered in the manifest above, one record per jurisdiction, so each
source keeps its own vintage, licence and `retrieved_at`. Manitoba is the one source that cannot be
refreshed automatically: it was provided **by request** in 2018 and Manitoba has published nothing
since, so it is the authoritative version rather than a stale copy --- but with no public endpoint,
a future revision will have to arrive the same way and cannot be version-checked from here.

# Derived input: the SCANFI land-cover reclassification {#sec-scanfi-lcc}

The caribou ranges are the input assembled from the most sources. The land-cover raster is the
input whose **record most easily misleads**, and it is worth documenting for the opposite reason:
it looks like a single published dataset, cites a DOI, and is not that dataset at all.

Until this revision the manifest carried one record, `scanfi-structure-2020`, covering "stand age,
biomass, land cover" and pointing at the SCANFI DOI under OGL. That is correct for age and biomass.
For land cover it was wrong on three counts at once: the file the pipeline consumes is a
**derived** reclassification rather than the published layer, the DOI does not resolve to it, and
the copy actually fetched is an access-controlled Drive object rather than an openly downloadable
one. Land cover is now registered separately as **`scanfi-lcc-2020`**, flagged `derived = TRUE`
with a `derived_from` pointer, so the manifest states what is really being read.

## Provenance chain

```
SCANFI_att_nfiLandcover_2020_v2_20260119.tif          published layer, 8 classes (1-8), Byte
  NRCan FTP (.tif + .qml/.aux.xml/.ovr sidecars)      <- what the DOI covers
        |
        |  LandR::convert_SCANFI_LCC_codes()          terra::subst(), 1:1 relabel
        v
SCANFI_att_nfiLandcover_CanadaLCCclassCodes_2020_v2_20260119.tif    <- what we consume
  Drive 1EGp7LUA7cXMR6KpXDmu617xsjwGM6aIx (2.73 GB, access-controlled)
        |
        |  + CA_FAO_forest_2019 (separate NFIS download)
        |    LCC[FAO == 2 & !LCC %in% c(210, 220, 230)] <- 240
        v
LCC with 240 "recently disturbed"                     LandWeb_preamble / prepInputs_SCANFI_LCC_FAO
```

The Drive object is a **precomputed convenience copy** so that runs do not repeat the reclass;
`LandR::convert_SCANFI_LCC_codes()` is what produced it, and can rebuild it (or the 1985--2025
siblings) from the published layer at any time. Nothing is lost by treating the FTP layer as the
source of record.

## The reclassification is a pure relabel

```{r}
#| label: tbl-scanfi-recode
#| tbl-cap: "LandR::convert_SCANFI_LCC_codes() is a single terra::subst(). Native class labels are from the published .qml sidecar; the LCC names are the NFI/EOSD legend LandR's LCC machinery is written against."
knitr::kable(
  data.table(
    Native = 1:8,
    `Native label` = c("Bryoid", "Herbs", "Rock", "Shrub", "TB (treed broadleaf)",
                       "TC (treed conifer)", "TM (treed mixed)", "Water"),
    `LCC code` = c(40L, 100L, 30L, 50L, 220L, 210L, 230L, 20L),
    `LCC name` = c("bryoids", "herbs", "snow_ice / rock_rubble / exposed_barren_land", "shrubs",
                   "broadleaf", "coniferous", "mixedwood", "water")
  ),
  align = "rlrl", booktabs = TRUE
) |>
  kable_styling(font_size = 8, latex_options = "hold_position")
```

It is a **bijection**: no class merges, splits, or disappears, and the thematic content is
untouched. Only the numbering changes, so that downstream LandR code written against the NFI/EOSD
code space (`overlayLCCs()`, `convertUnwantedLCC()`, `treeClassesLCC`) works unmodified. Two
consequences follow for anyone reading the legend comment in `LandWeb_preamble.R`, which lists the
full code space rather than the subset SCANFI can emit:

* **Wetland codes `80` / `81` are never produced.** SCANFI has no wetland class at all. Sampling
  four widely separated 1-degree windows (west-central Alberta, northern Yukon/NWT tundra, the
  Hudson Bay lowland, southern Saskatchewan prairie) returned nothing outside
  `{20, 30, 40, 50, 100, 210, 220, 230}`.
* **Code `240` is not in this file.** It is stamped downstream by the FAO overlay.

## Why treed wetlands arrive as forest

SCANFI classifies **what is growing on a pixel**, from structure resolved by imagery and NFI
photo-plots. NALCMS-style products --- including the 2020 Land Cover of Canada --- classify a
mixture of structure, landform and land use, so a single `Wetland` class absorbs everything from
open fen to black-spruce peatland regardless of canopy.

For LandWeb the structural reading is the one required. A treed wetland carries conifer biomass,
burns, and regenerates, so it must enter the vegetation and fire submodels as forest with a species
composition rather than be masked out as non-vegetated. The same argument applies to treed pixels
NALCMS assigns to cropland or shrubland. This is a deliberate property of choosing SCANFI, not an
artifact of the reclassification --- which, being a bijection, cannot introduce it.

The cost is that **wetness is invisible** in this input. Nothing downstream can separate an upland
black-spruce stand from a black-spruce peatland, though the two differ in productivity, fire
behaviour and succession. Recovering that distinction requires a separate wetland layer; it cannot
be recovered from this one.

### How much it moves, measured

Cross-tabulated against the 2020 Land Cover of Canada (NALCMS, EPSG:3979, reprojected onto the
SCANFI grid by nearest neighbour) over a 1-degree window in west-central Alberta, 13.2 M pixels:

```{r}
#| label: tbl-scanfi-nalcms
#| tbl-cap: "Where NALCMS 2020 classes land in the SCANFI legend, west-central Alberta window (13.2 M 30 m pixels)."
knitr::kable(
  data.table(
    `NALCMS 2020 class` = c("14 Wetland", "15 Cropland", "17 Urban", "18 Water"),
    `% of window` = c("5.96", "19.91", "2.55", "2.53"),
    `SCANFI calls it` = c("94% forested (210/220/230)", "75% herbs (100)",
                          "67% rock/barren (30)", "only 69% detected as water (20)")
  ),
  align = "lrl", booktabs = TRUE
) |>
  kable_styling(font_size = 8, latex_options = "hold_position")
```

Net effect in that window: **74.5% treed** under SCANFI against **54.3% forest** under NALCMS, a
~20 percentage-point difference sourced mostly from NALCMS wetland, cropland and shrubland.
Regionally the divergence is larger --- in the Hudson Bay lowland NALCMS reports 72.9% wetland
where SCANFI reports 45.6% coniferous, 32.5% shrub and 5.4% bryoid; on the prairie NALCMS's 73.1%
cropland is effectively renamed to 77.9% herbs. Conifer/broadleaf/mixed boundaries also genuinely
differ: SCANFI broadleaf is 20.9% of the Alberta window against 1.7% for the NALCMS broadleaf
class, with most SCANFI broadleaf falling in NALCMS *mixed forest*.

These are **not** errors in either product. They are two legends answering different questions, and
the provenance record should make clear which question this pipeline asked.

## Two format caveats

**The CRS carries no EPSG code.** It is Canada Atlas Lambert except that the latitude of false
origin is 0 rather than 49, so it is neither EPSG:3978 nor EPSG:3979 and tools will not recognise
it as either. Relative to EPSG:3979 it is a pure northing shift of 6,585,077 m (constant to
<0.1 m nationally; the ~0.5 m easting difference is NAD83 vs NAD83(CSRS) datum noise). That shift
is **not** a whole number of pixels --- 6585077 / 30 = 219502.57 --- so the two 30 m grids sit ~17 m
apart in northing and any reprojection resamples. Categorical data: use `method = "near"`.

**The precomputed copy is stored inefficiently.** The published layer is `Byte`, 256x256-tiled,
~1.5 GB. The precomputed reclassification is `Float32` written in full-width 178400x1 strips,
2.73 GB, for values that fit in a byte and have never exceeded 230. Any windowed read must pull
entire 178400-pixel rows at 4 bytes per pixel, which is part of why `LandWeb_preamble` performs its
own `terra::crop()` rather than letting `reproducible` warp the national raster.

This is fixed upstream rather than worked around here. `LandR::convert_SCANFI_LCC_codes()` now pins
`INT1U` with `NAflag = 255` and writes 256x256 LZW tiles (PredictiveEcology/LandR#200). All twelve
precomputed layers --- SCANFI v2 1985--2025 and v1 2000/2010/2020 --- were rebuilt from their
published sources and compared pixel by pixel against the copies currently distributed: **12 of 12
identical**, zero value mismatches and zero NA-pattern mismatches across each layer's
21,247,440,000 pixels, with per-class histograms equal. For 2020 those counts also reproduce the
histogram NRCan publishes in its own `.tif.aux.xml` sidecar, so rebuild and distributed copy agree
with the upstream product rather than merely with each other. Storage across the set falls from
32.98 GB to 15.05 GB. The rebuilt layers are drop-in replacements; the Drive copies are unchanged
pending a decision to swap them in.

## The FAO layer is a separate input

`CA_FAO_forest_2019` (NFIS; @WulderEtAl2020) is a distinct three-class raster --- `0` non-forest,
`1` current forest area, `2` temporally informed forest area --- where class `2` means forest *land
use* whose tree cover is temporarily absent after a stand-replacing disturbance. The recode rule
therefore reads: *FAO says this is forest land, SCANFI says nothing is standing on it, so mark it
recently disturbed (`240`)*. This is the only point at which FAO enters, and it happens after the
product described above; the land-cover file itself contains no FAO-derived values.

It also arrives on a different grid (`NAD_1983_Lambert_Conformal_Conic`, latitude of origin 49) and
must be cropped **before** reprojection, or `reproducible` warps the whole ~840 MB national raster
(>50 min) --- the bypass documented in `LandWeb_preamble.R`.

# Bibliography auto-sync

`workflowtools::sync_manifest_to_bibtex()` reads the manifest and writes a sidecar
`citations/data-sources.bib`, so that data-source citations do not have to be hand-maintained
alongside the methods references. It is wired as the `data_sources_bib` pipeline target and
deduplicates against the hand-maintained `citations/references.bib`.

**Adopted convention.** Data sources that are *also* cited as methods (SCANFI, Beaudoin *k*NN, the
NWAB reports) keep a single hand-maintained entry in `references.bib` and are marked `external` in
the manifest (so they are not duplicated into `data-sources.bib`). Any data source that needs a
citation but does *not* already have one in `references.bib` carries its BibTeX in the manifest
record and is emitted to `data-sources.bib` automatically. Reports list both `.bib` files in their
front matter, so either path resolves. This keeps the methods bibliography curated while making
data-only citations a by-product of registering the dataset --- no separate hand-maintenance.

# Per-report build-provenance appendix

Every report ends with a **Build provenance** appendix (see the appendix to this document, below),
generated at render time from `workflowtools`' `info_*()` family. It is a shared Quarto include ---
`reports/_provenance-appendix.qmd` --- that any report adopts with a single line
(`{{< include _provenance-appendix.qmd >}}`), so all reports are stamped identically. It records:

- **build identity** --- repository branch, HEAD commit, and render timestamp (`info_git()`);
- **pipeline state** --- the pinned SHA of every SpaDES module and owned package at render time
  (`info_submodules()`), the same revisions report 03 documents;
- **software environment** --- R version, operating system, and GEOS / GDAL / PROJ versions
  (`info_session()` / `info_spatial_libs()`).

For the simulation-run report (05) it will additionally carry the **base seed and per-replicate seed
derivation**; those apply once the main simulation is wired. (Note: `workflowtools`'
`reproducibility_receipt()` emits HTML `<details>` blocks aimed at HTML output, so the PDF reports
format the same `info_*()` data into tables directly.)

This appendix is what makes the v2-vs-v3 output comparison (report 04) interpretable: it pins each
PDF to the code and environment that produced it, so a reader can attribute an output difference to
code vs. inputs vs. randomness.

# Output provenance and lineage

Output provenance is the mirror of the input manifest: every artifact the pipeline produces --- a
summary CSV, a map, a rendered figure, a simulation raster --- should be traceable to **what
produced it**. The design has four ties per artifact:

1. **Producing target** --- the `targets` node and its command that wrote the file;
2. **Code revision** --- the SpaDES module / package SHAs in effect at build (the same
   `info_submodules()` set stamped in the build-provenance appendix);
3. **Input lineage** --- which registered inputs (`_input_manifest.json` ids) fed the target,
   transitively; and
4. **Run parameters** --- study area, replicate count, and random seed.

Much of this is **already captured, not yet surfaced**:

- **`targets` records it natively.** `targets::tar_meta()` stores, per target, the command, the
  upstream dependencies, a content hash, the seed, and the build timestamp --- i.e. the store *is* a
  content-addressed provenance graph. `targets::tar_network()` exposes the dependency edges (the
  input-to-output lineage) directly.
- **`SpaDES.targets` records per-run saves.** Each simulation stage returns an `outputs(sim)`
  manifest of the files it wrote (per-timestep saves, module `registerOutputs()` dumps, `Plots()`
  figures), discovered dynamically rather than declared a priori, plus a companion `format = "file"`
  target that hashes every saved file.

**What remains (planned):** a small producer that reduces `tar_meta()` + the run configuration into
an **output manifest** in the same shape as the input manifest (one record per artifact: id,
producing target, code SHA, input ids, run parameters, hash), and wiring the **build-provenance
appendix of a simulation report (05) to filter that manifest to the artifacts that report used** ---
closing the loop from a figure in a PDF back to the exact inputs, code, and seed behind it. This is
scoped to land with the main-simulation and summary stages (reports 04/05).

# Reproducibility standards

LandWeb v3 targets a **PERFICT**-style workflow [@McIntire:2022; @Barros2023] --- frequent **P**redictions and
**E**valuations of **R**eusable, **F**reely-accessible, **I**nteroperable models, built in
**C**ontinuous, **T**ested workflows --- and the manifest, appendix, and `targets` graph are the
runtime evidence for that claim. The table below maps the provenance design onto the standards the
project aims to meet; the "status" column is an honest self-assessment for the co-authors to refine.

| Standard | What it asks for | LandWeb v3 status |
|---|---|---|
| **FAIR** [@Wilkinson:2016] | Findable, Accessible, Interoperable, Reusable data + workflows | Inputs are registered with source, license, and citation (manifest); outputs FAIR-ify once the output manifest lands. Access to some inputs is controlled (proprietary/Drive). |
| **PERFICT** [@McIntire:2022; @Barros2023] | Reusable, tested, continuous predictive-ecology workflows | Core aim of the v3 rebuild: `targets` continuous workflow, owned CI-tested packages, per-FMA reuse. |
| **Ten Simple Rules** [@Sandve:2013] | Version control, recorded seeds, tracked intermediate steps, avoid manual steps | Met by `targets` (recorded seeds/hashes, no manual steps) + pinned revisions; NEWS now documents module history (report 03). |
| **ODD** [@Grimm:2006] | A standard narrative description of the model | Partly served by the manual (#128) + report 00; a formal ODD write-up is outstanding. |
| **TRACE** [@Grimm:2014] | Documented model development, testing, and analysis | Partly served by reports 01-03 (data, methods, changes) + module tests; a consolidated TRACE document is outstanding. |

The manifest + build-provenance appendix are the *runtime evidence* an ODD/TRACE narrative would
point to. Fleshing those narratives out (and the FLM-calibration best-practices alignment,
[@Reese:2024], report 07) is the next step for the co-authors.

\newpage

# References {.unnumbered}

::: {#refs}
:::

\newpage

# Provenance {.unnumbered}

- **Manifest:** `` `r mf` ``, built by `R/build_input_manifest.R` (hand-curated;
  `workflowtools::input_manifest_record()`), `r nrow(rec)` records.
- **Bibliography sync:** `citations/data-sources.bib`, generated by
  `workflowtools::sync_manifest_to_bibtex()` (the `data_sources_bib` target).
- **Method:** this report reads the committed manifest directly; rendered by the gated
  `landweb.extended_analyses` factory as a `tar_quarto()` target on the control node.

{{< include _provenance-appendix.qmd >}}
