# Running LandWeb

1. Launch Rstudio and open the LandWeb Rstudio project (`LandWeb.Rproj`);
2. Open the file `00-global.R` and run each line in sequence, responding to any prompts as required.

:::{.rmdwarning}
Before you can run the model, you first need to install the packages required for the project by restoring from the project's snapshot file.

```r
renv::restore()
```
:::

## Model setup and configuration

```r{setup-running-landweb, echo=FALSE, eval=TRUE, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, results = "hold") ## change to eval = TRUE if all chunks are to be evaluated

library("ggplot2")
library("sf")
```

The default settings for study area, model version, and scenario are defined in `01a-globalvars.R`.
These defaults are defined as 'dot-variables' (e.g., `.studyAreaName`) and can be set externally to the main script
(e.g., if `.studyAreaName` is defined before running `01a-globalvars.R`, then the user-set value will be used, rather than the default defined in that script).

### Select a study area

The model can be run over the entire study area, for certain individual provinces (currently only AB, SK, MB, NWT), or groups of predefined FMAs (see Fig. 1 and Table 1).

#### FMA boundaries

Currently, only a subset of the FMAs within the LandWeb study area are predefined to be run on their own (i.e., without needing to run the model over the entire study area; see Table 2).

\newpage
\blandscape

```{r fma-map, echo = FALSE, eval = 'true', fig.keep = 'all', out.width = '90%', results = 'hide', fig.cap = "\\label{fig:fma-map} FMAs within the LandWeb study area, identified by polygon IDs, which are described in Table 1. Polygon fill colours are used only to differentiate neighbouring FMAs."}
canProvs <- geodata::gadm(country = "CAN", level = 1, path = "../inputs") |>
  sf::st_as_sf()
lwProvs <- c("British Columbia", "Alberta", "Saskatchewan", "Manitoba")

fshp2 <- file.path("~/GitHub/LandWeb/inputs/FMA_Boundaries",
                   "FMA_Boundary_CustomProj/FMA_Boundary_Updated.shp")
sf2 <- sf::st_read(fshp2)
centroids_sf2 <- cbind(sf::st_centroid(sf2), sf::st_coordinates(sf::st_centroid(sf2$geometry)))

ggplot2::ggplot(data = sf2) +
  ggplot2::geom_sf(data = canProvs[canProvs$NAME_1 %in% lwProvs, ]) +
  ggplot2::geom_sf(mapping = ggplot2::aes(fill = as.factor(OBJECTID_1 %% 10))) +
  ggplot2::scale_fill_brewer(type = "div", palette = "Spectral") +
  ggplot2::geom_text(data = centroids_sf2, 
                     mapping = ggplot2::aes(x = X, y = Y, label = OBJECTID_1), size = 2) + 
  ggplot2::labs(title = "FMA boundaries within the LandWeb study area") + 
  ggplot2::xlab("Longitude") +
  ggplot2::ylab("Latitude") + 
  ggplot2::guides(fill = FALSE) +
  ggplot2::theme_bw()
```

\elandscape
\newpage

```{r fma-table, echo = FALSE, eval = TRUE}
data.frame(ID = sf2$OBJECTID_1, Name = sf2$Name) |>
  knitr::kable(caption = "FMA polygon IDs (from Figure 1) and their corresponding FMA names.",
               format = "latex", booktabs = TRUE, longtable = TRUE) |>
  kableExtra::kable_styling(latex_options =  c("striped", "repeat_header")) # font_size = 10
```

#### Choosing a study area

The model can be run on any of several pre-defined study areas summarized in the table below.
To select one of these predefined study areas, set `.studyAreaName` to use one of the following, corresponding to the polygon IDs in the map above.

```{r studyarea-table, echo = FALSE, eval = TRUE}
data.frame(
  studyAreaName = c("ANC", "Blueridge", "DMI", "Edson", "FMANWT ", "FMANWT2",
                    "LP_BC", "LP_MB", "Manning", "MillarWestern", "Mistik", "MPR",
                    "Sundre", "Tolko_AB_N", "Tolko_AB_S", "Tolko_SK", "Vanderwell",
                    "WestFraser_N", "WestFraser_S", "WeyCo_GP", "WeyCo_PT", "WeyCo_SK"),
  ID = c("39", "40", "42, 43", "53", "59", "58", "5, 6", "32", "45", "46", "15", "42, 43",
         "48", "50", "44, 49, 51", "22", "51, 52", "44, 51, 55", "53, 54", "56", "57", "21"), 
  Description = c(
    "ANC Timber Ltd.", "Blueridge Lumber Inc.",
    "Mercer Peace River Pulp Ltd. (formerly DMI)", 
    "West Fraser Mills Ltd. (Edson)", "Fort Resolution", "Fort Providence", 
    "Lousiana Pacific (British Columbia)", "Lousiana Pacific (Manitoba)", 
    "Manning Diversified Forest Products Ltd.", "Millar Western Forest Products Ltd.", 
    "Mistik", "Mercer Peace River Pulp Ltd. (formerly DMI)", "Sundre Forest Products Inc.", 
    "Tolko Industries Ltd. (Alberta North)", "Tolko Industries Ltd. (Alberta South)", 
    "Tolko (Saskatchewan)", "Vanderwell Contractors (1971) Ltd.", 
    "West Fraser Mills Ltd. (Slave Lake)", "West Fraser Mills Ltd. (Edson + Hinton)", 
    "Weyerhauser Company Ltd. (Grand Prairie)", "Weyerhauser Company Ltd. (Pembina Timberland)", 
    "Weyerhauser Company Ltd. (Pasquia-Porcupine)"
  )
) |>
  knitr::kable(caption = "Model study areas with corresponding FMA polygon IDs (from Figure 1).",
             format = "latex", booktabs = TRUE) |>
  kableExtra::kable_styling(latex_options = "striped") # font_size = 10
```

To run LandWeb over an entire province use one of `provAB`, `provMB`, `provNWT`, or `provSK`.

To run the entire LandWeb study area, use `LandWeb`.

\newpage

### Select a scenario

In version 2.0.0 of the LandWeb model, seed dispersal distances needed to be adjusted to ensure sufficient regeneration following fire.
These adjustments cause the model to behave more like a state-transition model, rather than a process-based one. 

Version 3.0.0 relaxes these parameter forcings to behave like the standard LANDIS-II model.

The dispersal scenario is principally set via `.version` but can be overridden is set via `.dispersalType`:

**Dispersal Scenario** | **Description**
-----------------------|----------------------------------------------------------------------------
`default` (v3)         | default LANDIS-II dispersal
`aspen`                | limit seed dispersal to deciduous only
`high` (v2)            | high seed dispersal of all species (used for v2.0.0 runs)
`none`                 | no seed dispersal (all species)

Additionally, v2.0.0 of the model uses adjusted fire return intervals (FRI) and log-adjusted rates of spread (ROS).
Version 3.0.0 uses the LandMine defaults.

The fire scenario is principally set via `.version` but can be overridden by specifying `.ROStype`:

**Fire Scenario** | **Description**
------------------|---------------------------------------------------------------------------------
`default` (v3)    | default rate of spread values
`burny`           | Increases the flammability of non-forest types to facilitate fire spread in landscapes with discontinuous fuels.
`equal`           | Set all rates of fire spread equal to each other (no vegetation differences)
`log` (v2)        | Reduce the rates of spread but keep the magnitude of vegetation differences

To define a scenario to run, select one dispersal scenario and one fire scenario from the tables above.
All LandWeb v2.0.0 runs from 2019 were run using `.dispersalType = "high"` and `.ROStype = "log"`.
Using v3.0.0 of the model, `.dispersalType = "default"` and `.ROStype = "default"`.

### Replication

To run multiple replicates of a given run, set `.rep` to an integer corresponding to the replicate id.
All replicate runs use a different random seed, and this seed is saved as on output for reuse in the event that a replicate needs to be rerun.
To rerun a replicate using a different seed, be sure to delete that run's `seed.rds` file.
The seed used is also saved in human-readable `seed.txt` file.

## Running the model

**NOTE:** The first time the model is run, it will automatically download additional data and install additional R packages, which can take some time to complete.

### Interactive R session

When working in an R session, be sure to set the working directory to the LandWeb project directory.
The first time running the model, open the file `00-global.R`, and step through each line to ensure any prompts etc. are answered correctly.

```r
source("00-global.R")
```

### Commandline interface

In addition to running the model in an interactive R session, we provide a commandline interface to run replicates of the model for the study areas defined above (i.e., batch mode).

For example, to run replicate number 7 of the model at $250 m$ resolution using FRI multiple of $1$ for Alberta FMU L11, use:

```bash
cd ~/GitHub/LandWeb

## ./run_fmu.sh <FMU> <FRI> <RES> <REP>
./run_fmu.sh L11 1 250 7
```

FMU commandline runs do not use modified dispersal nor fire scenarios, and thus only require the study area, replicate, and FRI multiple to be defined.

To run replicate number 7 of the model for the entire province of Alberta, use:

```bash
cd ~/GitHub/LandWeb

## ./run_fma.sh <FMU> <REP>
./run_fma.sh provAB 7
# ./run_fma_win.sh provAB 7 ## if on Windows!
```

FMA commandline runs use `highDispersal_logROS` scenarios at $250 m$ resolution, and thus only require the study area and replicate to be defined.

To run the entire LandWeb study area, only a replicate number needs to be passed.
For example:

```bash
## ./run_landweb.sh <REP>
./run_landweb.sh 7
```

LandWeb commandline run set the dispersal and fire scenarios as above, and the pixel resolution (`.pixelSize`) at $250 m$, and thus only require the study area and replicate to be defined.

## Changing model config values

See <https://github.com/PredictiveEcology/SpaDES.config/blob/development/R/config.landweb.R>

## Cache options

Simulation caching is provided by the `reproducible` and `SpaDES.core` packages, and is enabled by default.

The default cache uses a SQLite database backend; however, other backends can also be used.
More advanced users running multiple parallel simulations may wish to set up and use a PostgreSQL database for this cache.

## Post-processing analyses

After having run several reps of the model on a given study area, results are combined in subsequent post-processing analyses to generate the following outputs:

- boxplots of leading vegetation cover;
- histograms of leading vegetation cover;
- histograms of large patches.

To run processing, use `.mode = "postprocess"` and be sure to set the number of replicates run in the config.

## Additional Resources

Resources for (re)learning R and spatial data:
- <https://rspatial.org>
