---
title: "importLandWeb_BAP_Layers"
author: "Eliot McIntire and Alex M Chubaty"
date: "18 May 2018"
output:
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Overview

Provide an overview of what the module does / how to use the module.

Module documentation should be written so that others can use your module.
This is a template for module documentation, and should be changed to reflect your module.

## RMarkdown

RMarkdown syntax allows R code, outputs, and figures to be rendered in the documentation.

For help writing in RMarkdown, see http://rmarkdown.rstudio.com/.

# Usage

```{r module_usage}
library(igraph)
library(SpaDES)

moduleDir <- file.path("~/SpaDES/modules")
inputDir <- file.path(moduleDir, "inputs") %>% checkPath(create = TRUE)
outputDir <- file.path(moduleDir, "outputs") %>% checkPath(create = TRUE)
cacheDir <- file.path(outputDir, "cache") %>% checkPath(create = TRUE)
times <- list(start = 0, end = 10)
parameters <- list(
  #.progress = list(type = "text", interval = 1), # for a progress bar
  ## If there are further modules, each can have its own set of parameters:
  importLandWeb_BAP_Layers = list(.useCache = FALSE)
  #module2 = list(param1 = value1, param2 = value2)
)
modules <- list("importLandWeb_BAP_Layers")
objects <- list()
paths <- list(
  cachePath = cacheDir,
  modulePath = moduleDir,
  inputPath = inputDir,
  outputPath = outputDir
)

mySim <- simInit(times = times, params = parameters, modules = modules,
                 objects = objects, paths = paths)

spades(mySim)
```

# Events

Describe what happens for each event type.

## Plotting

Write what is plotted.

## Saving

Write what is saved.

# Data dependencies

## Input data

How to obtain input data, and a description of the data required by the module.
If `sourceURL` is specified, `downloadData("importLandWeb_BAP_Layers", "path/to/modules/dir")` may be sufficient.

## Output data

Description of the module outputs.

# Links to other modules

Describe any anticipated linkages to other modules.

# References

<!-- automatically generated; see https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html -->
